---
title: "Random forest"
format: html
---

# Setup  
```{r}
#| message: false
#| warning: false
library(doParallel)
library(tidymodels)
library(tidyverse)
library(vip)
library(sf)
library(partykit)
library(bonsai)
library(finetune)
```

# 1) M + W + S + O
# Yield
```{r }
df <- read_sf("../data/data_complete.geojson") %>% 
  st_drop_geometry() %>% 
  as.data.frame() %>% 
  dplyr::select(planting_date, digger_date, growing_season, variety, river_basin, soil_series, row_pattern, 
                irrigation, saved_seed, 
                tillage_method=tillage_method_simplify,
                rotation = rotation_simplify_years_wo_peanut, seeding_rate, 
                grade_n, 
                alt, stage_duration_emergence:elevation,
                lat= lat.y, lon = long
                ) %>% 
  mutate_if(is.character, as.factor) %>% 
  mutate(rotation = as.factor(rotation)) %>% 
  dplyr::select(-(c(alt, rotation, seeding_rate, tillage_method))) %>% 
  dplyr::select(!(c(soil_alpha_0_5:soil_bd_30_60, 
                    soil_hb_0_5:soil_hb_30_60, 
                    soil_lambda_0_5:soil_n_30_60,
                    soil_theta_r_0_5:soil_theta_r_30_60
                    ))) %>%
  dplyr::select(-(c(SCI:NGRDI))) %>% 
  dplyr::select(grade_n, planting_date:variety, 
                row_pattern:saved_seed, 
                lat, lon, 
                stage_duration_emergence:stage_vp_total,
                river_basin, soil_series,
                soil_clay_0_5:soil_theta_s_30_60,
                NDVI:elevation
                ) %>% 
  #dplyr::select(-(c(NDVI:GNDVI, elevation))) %>% 
  drop_na() %>% dplyr::filter(grade_n >65) 

df
```
# ML workflow  
## 1. Pre-processing  
### a. Data split  
For data split, let's use **70% training / 30% testing**.
```{r weather_split}
# Setting seed to get reproducible results  
set.seed(27)
weather_split <- initial_split(df, prop = 0.7, 
                               strata = "grade_n")
weather_split
```
```{r weather_train}
# Setting train set 
weather_train <- training(weather_split)

weather_train
```


```{r weather_test}
# Setting test split
weather_test <- testing(weather_split)

weather_test
```
```{r data distribution}
# Combine the data frames
plot_train <- weather_train %>% mutate(Dataset = "Train")
plot_test <- weather_test %>% mutate(Dataset = "Test")
combined_data <- bind_rows(plot_train, plot_test)

# Create the density plot
ggplot(data = combined_data) +
  geom_density(aes(x = grade_n, fill = Dataset), alpha = 0.5) +
  scale_fill_manual(values = c("red", "blue")) +
  labs(x = "Yield (kg/ha)", y = "Density", fill = "Dataset")
```
  
### b. Data processing  
```{r weather_recipe}
weather_recipe <-
  # Defining predicted and predictor variables
  recipe(grade_n ~ .,
         data = weather_train) 
weather_recipe
```

```{r weather_prep}
weather_prep <- weather_recipe %>%
  prep()

weather_prep
```
## 2. Training  
### a. Model specification  
```{r hyperparameters}
cart_spec <- 
  decision_tree(tree_depth = tune(),
                min_n = tune()
                ) %>% 
  # Specify the engine
  set_engine("partykit", 
             conditional_min_criterion = tune()
             ) %>% 
  # Specifying mode  
  set_mode("regression")

cart_spec
```

### b. Hyper-parameter tuning  
  
```{r resampling_foldcv}
set.seed(15)
resampling_foldcv <- vfold_cv(weather_train, 
                              v = 5)

resampling_foldcv
```
```{r cit_param}
cit_param <- cart_spec %>% 
  extract_parameter_set_dials() %>% 
  update(conditional_min_criterion =conditional_min_criterion())

cit_param
cit_param$object
```

```{r rf_grid_result}
set.seed(45)
registerDoParallel(cores = parallel::detectCores() - 1)
rf_grid_result <- tune_race_anova(object = cart_spec,
                                   preprocessor = weather_recipe,
                                   resamples = resampling_foldcv,
                                   param_info = cit_param,
                                  grid = 50
                      )
stopImplicitCluster()
rf_grid_result$.metrics[[5]]
autoplot(rf_grid_result)
beepr::beep()
```

```{r hyperparameters selection by RMSE}
# Based on lowest RMSE
best_rmse <- rf_grid_result %>% 
  select_by_pct_loss("tree_depth",
                     metric = "rmse",
                     limit = 2
                     )%>% 
  mutate(source = "best_rmse")

best_rmse_2 <- rf_grid_result %>% 
  select_best(metric = "rmse")%>% 
  mutate(source = "best_rmse_2")

best_rmse_3 <- rf_grid_result %>% 
  select_by_one_std_err(metric = "rmse",
                        tree_depth
                        )%>% 
  mutate(source = "best_rmse_3")

```

```{r hyperparameters selection by R2}
# Based on greatest R2
best_r2 <- rf_grid_result %>% 
  select_by_pct_loss("tree_depth",
                     metric = "rsq",
                     limit = 2
                     ) %>% 
  mutate(source = "best_r2")

best_r2_2 <- rf_grid_result %>% 
  select_best(metric = "rsq")%>% 
  mutate(source = "best_r2_2")

best_r2_3 <- rf_grid_result %>% 
  select_by_one_std_err(metric = "rsq",
                        tree_depth
                        ) %>%
  mutate(source = "best_r2_3")
```
```{r comparing values}
hyperparameters_df <- best_rmse %>% 
  bind_rows(best_rmse_2, best_rmse_3, best_r2, best_r2_2, best_r2_3)
```

```{r}
# Function to find the best model
compare_hyperparameters <- function(params_df, recipe, split) {
  # Create empty tibble to store results
  results <- tibble(
    tree_depth = numeric(),
    min_n = numeric(),
    conditional_min_criterion = numeric(),
    rmse = numeric(),
    source = character()
  )
  
  # Loop through each row of parameters
  for(i in 1:nrow(params_df)) {
    # Extract current parameters
    current_params <- params_df[i, ]
    
    # Create model specification
    set.seed(10)
    current_spec <- decision_tree(
      tree_depth = current_params$tree_depth,
      min_n = current_params$min_n
      ) %>%
      set_engine('partykit',
      conditional_min_criterion = current_params$current_params
                 ) %>%
      set_mode('regression')
    
    # Fit model and collect metrics
    current_fit <- last_fit(current_spec, recipe, split = split)
    current_metrics <- current_fit %>% collect_metrics()
    
    # Extract RMSE value
    current_rmse <- current_metrics %>% 
      filter(.metric == "rmse") %>% 
      pull(.estimate)
    
    # Store results
    results <- results %>% add_row(
      tree_depth = current_params$tree_depth,
      min_n = current_params$min_n,
      conditional_min_criterion =current_params$conditional_min_criterion,
      rmse = current_rmse,
      source = current_params$source
    )
  }
  
  # Find the best combination
  best_params <- results %>% 
    arrange(rmse) %>% 
    slice(1)
  
  return(list(
    all_results = results,
    best_combination = best_params
  ))
}

# Example usage
results <- compare_hyperparameters(
  params_df = hyperparameters_df,
  recipe = weather_recipe,
  split = weather_split
)
print(results)
```




```{r final_spec}
final_spec <- decision_tree(tree_depth = best_rmse$tree_depth,
                            min_n = best_rmse$min_n
                            ) %>%
  # Specify the engine
  set_engine("partykit",
             conditional_min_criterion = best_rmse$conditional_min_criterion) %>%
  # Specifying mode  
  set_mode("regression")

final_spec
```

## 3. Validation  
```{r final_fit}
set.seed(10)
final_fit <- last_fit(final_spec,
                weather_recipe,
                split = weather_split)

final_fit %>%
  collect_predictions()
```
Metrics on the **test set**:
```{r}
final_fit %>%
  collect_metrics()
```
Metrics on **train set** (for curiosity and compare to test set):  
```{r}
# RMSE
final_spec %>%
  fit(grade_n ~ .,
      data = bake(weather_prep, 
                  weather_train)) %>%
  augment(new_data = bake(weather_prep, 
                          weather_train)) %>% 
  rmse(grade_n, .pred) %>%
  bind_rows(
    
    
    # R2
    final_spec %>%
      fit(grade_n ~ .,
          data = bake(weather_prep, 
                      weather_train)) %>%
      augment(new_data = bake(weather_prep, 
                              weather_train)) %>% 
      rsq(grade_n, .pred)
    
  )

```
```{r predicted vs. observed plot}
plot <- final_fit %>%
  collect_predictions() %>%
  metrica::scatter_plot(obs = grade_n,
                        pred = .pred,
                        print_eq = T,
                        print_metrics = T,
                        metrics_list = c("R2", "RMSE"),
                      # Customize metrics position
                        position_metrics = c(x=75,y=69),
                        position_eq = c(x=75, y =70.5))

plot
```


```{r export first plot}
ggsave(plot = plot,
       filename = "../output/ML_Grade/15_CART/M+W+S+O.png",
       width = 5,
       height = 5
       )
```

### Variable importance: 

```{r plotting VI}
set.seed(10)
plot_2 <- final_spec %>%
  fit(grade_n ~ .,
         data = bake(weather_prep, df)) %>%
    vip::vi() %>% 
  mutate(
    Importance = abs(Importance),
    Variable = fct_reorder(Variable, 
                           Importance)
  ) %>%
  slice_head(n = 10) %>%
  ggplot(aes(x = Importance, 
             y = Variable)) +
  geom_col() +
  scale_x_continuous(expand = c(0, 0)) +
  labs(y = NULL)

plot_2
```


```{r export second plot}
ggsave(plot = plot_2,
       filename = "../output/ML_Grade/15_CART/M+W+S+O_2.png",
       width = 5,
       height = 5,
       bg = "white"
       ) 
```

# 2) M + W + S 
# Yield
```{r }
df <- read_sf("../data/data_complete.geojson") %>% 
  st_drop_geometry() %>% 
  as.data.frame() %>% 
  dplyr::select(planting_date, digger_date, growing_season, variety, river_basin, soil_series, row_pattern, 
                irrigation, saved_seed, 
                tillage_method=tillage_method_simplify,
                rotation = rotation_simplify_years_wo_peanut, seeding_rate, 
                grade_n, 
                alt, stage_duration_emergence:elevation,
                lat= lat.y, lon = long
                ) %>% 
  mutate_if(is.character, as.factor) %>% 
  mutate(rotation = as.factor(rotation)) %>% 
  dplyr::select(-(c(alt, rotation, seeding_rate, tillage_method))) %>% 
  dplyr::select(!(c(soil_alpha_0_5:soil_bd_30_60, 
                    soil_hb_0_5:soil_hb_30_60, 
                    soil_lambda_0_5:soil_n_30_60,
                    soil_theta_r_0_5:soil_theta_r_30_60
                    ))) %>%
  dplyr::select(-(c(SCI:NGRDI))) %>% 
  dplyr::select(grade_n, planting_date:variety, 
                row_pattern:saved_seed, 
                lat, lon, 
                stage_duration_emergence:stage_vp_total,
                river_basin, soil_series,
                soil_clay_0_5:soil_theta_s_30_60#,
                #NDVI:elevation
                ) %>% 
  #dplyr::select(-(c(NDVI:GNDVI, elevation))) %>% 
  drop_na() %>% dplyr::filter(grade_n >65) 

df
```
# ML workflow  
## 1. Pre-processing  
### a. Data split  
For data split, let's use **70% training / 30% testing**.
```{r weather_split}
# Setting seed to get reproducible results  
set.seed(27)
weather_split <- initial_split(df, prop = 0.7, 
                               strata = "grade_n")
weather_split
```
```{r weather_train}
# Setting train set 
weather_train <- training(weather_split)

weather_train
```


```{r weather_test}
# Setting test split
weather_test <- testing(weather_split)

weather_test
```
```{r data distribution}
# Combine the data frames
plot_train <- weather_train %>% mutate(Dataset = "Train")
plot_test <- weather_test %>% mutate(Dataset = "Test")
combined_data <- bind_rows(plot_train, plot_test)

# Create the density plot
ggplot(data = combined_data) +
  geom_density(aes(x = grade_n, fill = Dataset), alpha = 0.5) +
  scale_fill_manual(values = c("red", "blue")) +
  labs(x = "Yield (kg/ha)", y = "Density", fill = "Dataset")
```
  
### b. Data processing  
```{r weather_recipe}
weather_recipe <-
  # Defining predicted and predictor variables
  recipe(grade_n ~ .,
         data = weather_train) 
weather_recipe
```

```{r weather_prep}
weather_prep <- weather_recipe %>%
  prep()

weather_prep
```
## 2. Training  
### a. Model specification  
```{r hyperparameters}
cart_spec <- 
  decision_tree(tree_depth = tune(),
                min_n = tune()
                ) %>% 
  # Specify the engine
  set_engine("partykit", 
             conditional_min_criterion = tune()
             ) %>% 
  # Specifying mode  
  set_mode("regression")

cart_spec
```

### b. Hyper-parameter tuning  
  
```{r resampling_foldcv}
set.seed(15)
resampling_foldcv <- vfold_cv(weather_train, 
                              v = 5)

resampling_foldcv
```
```{r cit_param}
cit_param <- cart_spec %>% 
  extract_parameter_set_dials() %>% 
  update(conditional_min_criterion =conditional_min_criterion())

cit_param
cit_param$object
```

```{r rf_grid_result}
set.seed(45)
registerDoParallel(cores = parallel::detectCores() - 1)
rf_grid_result <- tune_race_anova(object = cart_spec,
                                   preprocessor = weather_recipe,
                                   resamples = resampling_foldcv,
                                   param_info = cit_param,
                                  grid = 50
                      )
stopImplicitCluster()
rf_grid_result$.metrics[[5]]
autoplot(rf_grid_result)
beepr::beep()
```

```{r hyperparameters selection by RMSE}
# Based on lowest RMSE
best_rmse <- rf_grid_result %>% 
  select_by_pct_loss("tree_depth",
                     metric = "rmse",
                     limit = 2
                     )%>% 
  mutate(source = "best_rmse")

best_rmse_2 <- rf_grid_result %>% 
  select_best(metric = "rmse")%>% 
  mutate(source = "best_rmse_2")

best_rmse_3 <- rf_grid_result %>% 
  select_by_one_std_err(metric = "rmse",
                        tree_depth
                        )%>% 
  mutate(source = "best_rmse_3")

```

```{r hyperparameters selection by R2}
# Based on greatest R2
best_r2 <- rf_grid_result %>% 
  select_by_pct_loss("tree_depth",
                     metric = "rsq",
                     limit = 2
                     ) %>% 
  mutate(source = "best_r2")

best_r2_2 <- rf_grid_result %>% 
  select_best(metric = "rsq")%>% 
  mutate(source = "best_r2_2")

best_r2_3 <- rf_grid_result %>% 
  select_by_one_std_err(metric = "rsq",
                        tree_depth
                        ) %>%
  mutate(source = "best_r2_3")
```
```{r comparing values}
hyperparameters_df <- best_rmse %>% 
  bind_rows(best_rmse_2, best_rmse_3, best_r2, best_r2_2, best_r2_3)
```

```{r}
# Function to find the best model
compare_hyperparameters <- function(params_df, recipe, split) {
  # Create empty tibble to store results
  results <- tibble(
    tree_depth = numeric(),
    min_n = numeric(),
    conditional_min_criterion = numeric(),
    rmse = numeric(),
    source = character()
  )
  
  # Loop through each row of parameters
  for(i in 1:nrow(params_df)) {
    # Extract current parameters
    current_params <- params_df[i, ]
    
    # Create model specification
    set.seed(10)
    current_spec <- decision_tree(
      tree_depth = current_params$tree_depth,
      min_n = current_params$min_n
      ) %>%
      set_engine('partykit',
      conditional_min_criterion = current_params$current_params
                 ) %>%
      set_mode('regression')
    
    # Fit model and collect metrics
    current_fit <- last_fit(current_spec, recipe, split = split)
    current_metrics <- current_fit %>% collect_metrics()
    
    # Extract RMSE value
    current_rmse <- current_metrics %>% 
      filter(.metric == "rmse") %>% 
      pull(.estimate)
    
    # Store results
    results <- results %>% add_row(
      tree_depth = current_params$tree_depth,
      min_n = current_params$min_n,
      conditional_min_criterion =current_params$conditional_min_criterion,
      rmse = current_rmse,
      source = current_params$source
    )
  }
  
  # Find the best combination
  best_params <- results %>% 
    arrange(rmse) %>% 
    slice(1)
  
  return(list(
    all_results = results,
    best_combination = best_params
  ))
}

# Example usage
results <- compare_hyperparameters(
  params_df = hyperparameters_df,
  recipe = weather_recipe,
  split = weather_split
)
print(results)
```




```{r final_spec}
final_spec <- decision_tree(tree_depth = best_rmse$tree_depth,
                            min_n = best_rmse$min_n
                            ) %>%
  # Specify the engine
  set_engine("partykit",
             conditional_min_criterion = best_rmse$conditional_min_criterion) %>%
  # Specifying mode  
  set_mode("regression")

final_spec
```

## 3. Validation  
```{r final_fit}
set.seed(10)
final_fit <- last_fit(final_spec,
                weather_recipe,
                split = weather_split)

final_fit %>%
  collect_predictions()
```
Metrics on the **test set**:
```{r}
final_fit %>%
  collect_metrics()
```
Metrics on **train set** (for curiosity and compare to test set):  
```{r}
# RMSE
final_spec %>%
  fit(grade_n ~ .,
      data = bake(weather_prep, 
                  weather_train)) %>%
  augment(new_data = bake(weather_prep, 
                          weather_train)) %>% 
  rmse(grade_n, .pred) %>%
  bind_rows(
    
    
    # R2
    final_spec %>%
      fit(grade_n ~ .,
          data = bake(weather_prep, 
                      weather_train)) %>%
      augment(new_data = bake(weather_prep, 
                              weather_train)) %>% 
      rsq(grade_n, .pred)
    
  )

```
```{r predicted vs. observed plot}
plot <- final_fit %>%
  collect_predictions() %>%
  metrica::scatter_plot(obs = grade_n,
                        pred = .pred,
                        print_eq = T,
                        print_metrics = T,
                        metrics_list = c("R2", "RMSE"),
                      # Customize metrics position
                        position_metrics = c(x=75,y=71),
                        position_eq = c(x=75, y =72.5))

plot
```


```{r export first plot}
ggsave(plot = plot,
       filename = "../output/ML_Grade/15_CART/M+W+S.png",
       width = 5,
       height = 5
       )
```

### Variable importance: 

```{r plotting VI}
set.seed(10)
plot_2 <- final_spec %>%
  fit(grade_n ~ .,
         data = bake(weather_prep, df)) %>%
    vip::vi() %>% 
  mutate(
    Importance = abs(Importance),
    Variable = fct_reorder(Variable, 
                           Importance)
  ) %>%
  slice_head(n = 10) %>%
  ggplot(aes(x = Importance, 
             y = Variable)) +
  geom_col() +
  scale_x_continuous(expand = c(0, 0)) +
  labs(y = NULL)

plot_2
```


```{r export second plot}
ggsave(plot = plot_2,
       filename = "../output/ML_Grade/15_CART/M+W+S_2.png",
       width = 5,
       height = 5,
       bg = "white"
       ) 
```

# 3) M + W + O
# Yield
```{r }
df <- read_sf("../data/data_complete.geojson") %>% 
  st_drop_geometry() %>% 
  as.data.frame() %>% 
  dplyr::select(planting_date, digger_date, growing_season, variety, river_basin, soil_series, row_pattern, 
                irrigation, saved_seed, 
                tillage_method=tillage_method_simplify,
                rotation = rotation_simplify_years_wo_peanut, seeding_rate, 
                grade_n, 
                alt, stage_duration_emergence:elevation,
                lat= lat.y, lon = long
                ) %>% 
  mutate_if(is.character, as.factor) %>% 
  mutate(rotation = as.factor(rotation)) %>% 
  dplyr::select(-(c(alt, rotation, seeding_rate, tillage_method))) %>% 
  dplyr::select(!(c(soil_alpha_0_5:soil_bd_30_60, 
                    soil_hb_0_5:soil_hb_30_60, 
                    soil_lambda_0_5:soil_n_30_60,
                    soil_theta_r_0_5:soil_theta_r_30_60
                    ))) %>%
  dplyr::select(-(c(SCI:NGRDI))) %>% 
  dplyr::select(grade_n, planting_date:variety, 
                row_pattern:saved_seed, 
                lat, lon, 
                stage_duration_emergence:stage_vp_total,
                #river_basin, soil_series,
                #soil_clay_0_5:soil_theta_s_30_60,
                NDVI:elevation
                ) %>% 
  #dplyr::select(-(c(NDVI:GNDVI, elevation))) %>% 
  drop_na() %>% dplyr::filter(grade_n >65) 

df
```
# ML workflow  
## 1. Pre-processing  
### a. Data split  
For data split, let's use **70% training / 30% testing**.
```{r weather_split}
# Setting seed to get reproducible results  
set.seed(27)
weather_split <- initial_split(df, prop = 0.7, 
                               strata = "grade_n")
weather_split
```
```{r weather_train}
# Setting train set 
weather_train <- training(weather_split)

weather_train
```


```{r weather_test}
# Setting test split
weather_test <- testing(weather_split)

weather_test
```
```{r data distribution}
# Combine the data frames
plot_train <- weather_train %>% mutate(Dataset = "Train")
plot_test <- weather_test %>% mutate(Dataset = "Test")
combined_data <- bind_rows(plot_train, plot_test)

# Create the density plot
ggplot(data = combined_data) +
  geom_density(aes(x = grade_n, fill = Dataset), alpha = 0.5) +
  scale_fill_manual(values = c("red", "blue")) +
  labs(x = "Yield (kg/ha)", y = "Density", fill = "Dataset")
```
  
### b. Data processing  
```{r weather_recipe}
weather_recipe <-
  # Defining predicted and predictor variables
  recipe(grade_n ~ .,
         data = weather_train) 
weather_recipe
```

```{r weather_prep}
weather_prep <- weather_recipe %>%
  prep()

weather_prep
```
## 2. Training  
### a. Model specification  
```{r hyperparameters}
cart_spec <- 
  decision_tree(tree_depth = tune(),
                min_n = tune()
                ) %>% 
  # Specify the engine
  set_engine("partykit", 
             conditional_min_criterion = tune()
             ) %>% 
  # Specifying mode  
  set_mode("regression")

cart_spec
```

### b. Hyper-parameter tuning  
  
```{r resampling_foldcv}
set.seed(15)
resampling_foldcv <- vfold_cv(weather_train, 
                              v = 5)

resampling_foldcv
```
```{r cit_param}
cit_param <- cart_spec %>% 
  extract_parameter_set_dials() %>% 
  update(conditional_min_criterion =conditional_min_criterion())

cit_param
cit_param$object
```

```{r rf_grid_result}
set.seed(45)
registerDoParallel(cores = parallel::detectCores() - 1)
rf_grid_result <- tune_race_anova(object = cart_spec,
                                   preprocessor = weather_recipe,
                                   resamples = resampling_foldcv,
                                   param_info = cit_param,
                                  grid = 50
                      )
stopImplicitCluster()
rf_grid_result$.metrics[[5]]
autoplot(rf_grid_result)
beepr::beep()
```

```{r hyperparameters selection by RMSE}
# Based on lowest RMSE
best_rmse <- rf_grid_result %>% 
  select_by_pct_loss("tree_depth",
                     metric = "rmse",
                     limit = 2
                     )%>% 
  mutate(source = "best_rmse")

best_rmse_2 <- rf_grid_result %>% 
  select_best(metric = "rmse")%>% 
  mutate(source = "best_rmse_2")

best_rmse_3 <- rf_grid_result %>% 
  select_by_one_std_err(metric = "rmse",
                        tree_depth
                        )%>% 
  mutate(source = "best_rmse_3")

```

```{r hyperparameters selection by R2}
# Based on greatest R2
best_r2 <- rf_grid_result %>% 
  select_by_pct_loss("tree_depth",
                     metric = "rsq",
                     limit = 2
                     ) %>% 
  mutate(source = "best_r2")

best_r2_2 <- rf_grid_result %>% 
  select_best(metric = "rsq")%>% 
  mutate(source = "best_r2_2")

best_r2_3 <- rf_grid_result %>% 
  select_by_one_std_err(metric = "rsq",
                        tree_depth
                        ) %>%
  mutate(source = "best_r2_3")
```
```{r comparing values}
hyperparameters_df <- best_rmse %>% 
  bind_rows(best_rmse_2, best_rmse_3, best_r2, best_r2_2, best_r2_3)
```

```{r}
# Function to find the best model
compare_hyperparameters <- function(params_df, recipe, split) {
  # Create empty tibble to store results
  results <- tibble(
    tree_depth = numeric(),
    min_n = numeric(),
    conditional_min_criterion = numeric(),
    rmse = numeric(),
    source = character()
  )
  
  # Loop through each row of parameters
  for(i in 1:nrow(params_df)) {
    # Extract current parameters
    current_params <- params_df[i, ]
    
    # Create model specification
    set.seed(10)
    current_spec <- decision_tree(
      tree_depth = current_params$tree_depth,
      min_n = current_params$min_n
      ) %>%
      set_engine('partykit',
      conditional_min_criterion = current_params$current_params
                 ) %>%
      set_mode('regression')
    
    # Fit model and collect metrics
    current_fit <- last_fit(current_spec, recipe, split = split)
    current_metrics <- current_fit %>% collect_metrics()
    
    # Extract RMSE value
    current_rmse <- current_metrics %>% 
      filter(.metric == "rmse") %>% 
      pull(.estimate)
    
    # Store results
    results <- results %>% add_row(
      tree_depth = current_params$tree_depth,
      min_n = current_params$min_n,
      conditional_min_criterion =current_params$conditional_min_criterion,
      rmse = current_rmse,
      source = current_params$source
    )
  }
  
  # Find the best combination
  best_params <- results %>% 
    arrange(rmse) %>% 
    slice(1)
  
  return(list(
    all_results = results,
    best_combination = best_params
  ))
}

# Example usage
results <- compare_hyperparameters(
  params_df = hyperparameters_df,
  recipe = weather_recipe,
  split = weather_split
)
print(results)
```




```{r final_spec}
final_spec <- decision_tree(tree_depth = best_rmse$tree_depth,
                            min_n = best_rmse$min_n
                            ) %>%
  # Specify the engine
  set_engine("partykit",
             conditional_min_criterion = best_rmse$conditional_min_criterion) %>%
  # Specifying mode  
  set_mode("regression")

final_spec
```

## 3. Validation  
```{r final_fit}
set.seed(10)
final_fit <- last_fit(final_spec,
                weather_recipe,
                split = weather_split)

final_fit %>%
  collect_predictions()
```
Metrics on the **test set**:
```{r}
final_fit %>%
  collect_metrics()
```
Metrics on **train set** (for curiosity and compare to test set):  
```{r}
# RMSE
final_spec %>%
  fit(grade_n ~ .,
      data = bake(weather_prep, 
                  weather_train)) %>%
  augment(new_data = bake(weather_prep, 
                          weather_train)) %>% 
  rmse(grade_n, .pred) %>%
  bind_rows(
    
    
    # R2
    final_spec %>%
      fit(grade_n ~ .,
          data = bake(weather_prep, 
                      weather_train)) %>%
      augment(new_data = bake(weather_prep, 
                              weather_train)) %>% 
      rsq(grade_n, .pred)
    
  )

```
```{r predicted vs. observed plot}
plot <- final_fit %>%
  collect_predictions() %>%
  metrica::scatter_plot(obs = grade_n,
                        pred = .pred,
                        print_eq = T,
                        print_metrics = T,
                        metrics_list = c("R2", "RMSE"),
                      # Customize metrics position
                        position_metrics = c(x=75,y=73),
                        position_eq = c(x=75, y =71))

plot
```


```{r export first plot}
ggsave(plot = plot,
       filename = "../output/ML_Grade/15_CART/M+W+O.png",
       width = 5,
       height = 5
       )
```

### Variable importance: 

```{r plotting VI}
set.seed(10)
plot_2 <- final_spec %>%
  fit(grade_n ~ .,
         data = bake(weather_prep, df)) %>%
    vip::vi() %>% 
  mutate(
    Importance = abs(Importance),
    Variable = fct_reorder(Variable, 
                           Importance)
  ) %>%
  slice_head(n = 10) %>%
  ggplot(aes(x = Importance, 
             y = Variable)) +
  geom_col() +
  scale_x_continuous(expand = c(0, 0)) +
  labs(y = NULL)

plot_2
```


```{r export second plot}
ggsave(plot = plot_2,
       filename = "../output/ML_Grade/15_CART/M+W+O_2.png",
       width = 5,
       height = 5,
       bg = "white"
       ) 
```

# 4) M + S + O
# Yield
```{r }
df <- read_sf("../data/data_complete.geojson") %>% 
  st_drop_geometry() %>% 
  as.data.frame() %>% 
  dplyr::select(planting_date, digger_date, growing_season, variety, river_basin, soil_series, row_pattern, 
                irrigation, saved_seed, 
                tillage_method=tillage_method_simplify,
                rotation = rotation_simplify_years_wo_peanut, seeding_rate, 
                grade_n, 
                alt, stage_duration_emergence:elevation,
                lat= lat.y, lon = long
                ) %>% 
  mutate_if(is.character, as.factor) %>% 
  mutate(rotation = as.factor(rotation)) %>% 
  dplyr::select(-(c(alt, rotation, seeding_rate, tillage_method))) %>% 
  dplyr::select(!(c(soil_alpha_0_5:soil_bd_30_60, 
                    soil_hb_0_5:soil_hb_30_60, 
                    soil_lambda_0_5:soil_n_30_60,
                    soil_theta_r_0_5:soil_theta_r_30_60
                    ))) %>%
  dplyr::select(-(c(SCI:NGRDI))) %>% 
  dplyr::select(grade_n, planting_date:variety, 
                row_pattern:saved_seed, 
                lat, lon, 
                #stage_duration_emergence:stage_vp_total,
                river_basin, soil_series,
                soil_clay_0_5:soil_theta_s_30_60,
                NDVI:elevation
                ) %>% 
  #dplyr::select(-(c(NDVI:GNDVI, elevation))) %>% 
  drop_na() %>% dplyr::filter(grade_n >65) 

df
```
# ML workflow  
## 1. Pre-processing  
### a. Data split  
For data split, let's use **70% training / 30% testing**.
```{r weather_split}
# Setting seed to get reproducible results  
set.seed(27)
weather_split <- initial_split(df, prop = 0.7, 
                               strata = "grade_n")
weather_split
```
```{r weather_train}
# Setting train set 
weather_train <- training(weather_split)

weather_train
```


```{r weather_test}
# Setting test split
weather_test <- testing(weather_split)

weather_test
```
```{r data distribution}
# Combine the data frames
plot_train <- weather_train %>% mutate(Dataset = "Train")
plot_test <- weather_test %>% mutate(Dataset = "Test")
combined_data <- bind_rows(plot_train, plot_test)

# Create the density plot
ggplot(data = combined_data) +
  geom_density(aes(x = grade_n, fill = Dataset), alpha = 0.5) +
  scale_fill_manual(values = c("red", "blue")) +
  labs(x = "Yield (kg/ha)", y = "Density", fill = "Dataset")
```
  
### b. Data processing  
```{r weather_recipe}
weather_recipe <-
  # Defining predicted and predictor variables
  recipe(grade_n ~ .,
         data = weather_train) 
weather_recipe
```

```{r weather_prep}
weather_prep <- weather_recipe %>%
  prep()

weather_prep
```
## 2. Training  
### a. Model specification  
```{r hyperparameters}
cart_spec <- 
  decision_tree(tree_depth = tune(),
                min_n = tune()
                ) %>% 
  # Specify the engine
  set_engine("partykit", 
             conditional_min_criterion = tune()
             ) %>% 
  # Specifying mode  
  set_mode("regression")

cart_spec
```

### b. Hyper-parameter tuning  
  
```{r resampling_foldcv}
set.seed(15)
resampling_foldcv <- vfold_cv(weather_train, 
                              v = 5)

resampling_foldcv
```
```{r cit_param}
cit_param <- cart_spec %>% 
  extract_parameter_set_dials() %>% 
  update(conditional_min_criterion =conditional_min_criterion())

cit_param
cit_param$object
```

```{r rf_grid_result}
set.seed(45)
registerDoParallel(cores = parallel::detectCores() - 1)
rf_grid_result <- tune_race_anova(object = cart_spec,
                                   preprocessor = weather_recipe,
                                   resamples = resampling_foldcv,
                                   param_info = cit_param,
                                  grid = 50
                      )
stopImplicitCluster()
rf_grid_result$.metrics[[5]]
autoplot(rf_grid_result)
beepr::beep()
```

```{r hyperparameters selection by RMSE}
# Based on lowest RMSE
best_rmse <- rf_grid_result %>% 
  select_by_pct_loss("tree_depth",
                     metric = "rmse",
                     limit = 2
                     )%>% 
  mutate(source = "best_rmse")

best_rmse_2 <- rf_grid_result %>% 
  select_best(metric = "rmse")%>% 
  mutate(source = "best_rmse_2")

best_rmse_3 <- rf_grid_result %>% 
  select_by_one_std_err(metric = "rmse",
                        tree_depth
                        )%>% 
  mutate(source = "best_rmse_3")

```

```{r hyperparameters selection by R2}
# Based on greatest R2
best_r2 <- rf_grid_result %>% 
  select_by_pct_loss("tree_depth",
                     metric = "rsq",
                     limit = 2
                     ) %>% 
  mutate(source = "best_r2")

best_r2_2 <- rf_grid_result %>% 
  select_best(metric = "rsq")%>% 
  mutate(source = "best_r2_2")

best_r2_3 <- rf_grid_result %>% 
  select_by_one_std_err(metric = "rsq",
                        tree_depth
                        ) %>%
  mutate(source = "best_r2_3")
```
```{r comparing values}
hyperparameters_df <- best_rmse %>% 
  bind_rows(best_rmse_2, best_rmse_3, best_r2, best_r2_2, best_r2_3)
```

```{r}
# Function to find the best model
compare_hyperparameters <- function(params_df, recipe, split) {
  # Create empty tibble to store results
  results <- tibble(
    tree_depth = numeric(),
    min_n = numeric(),
    conditional_min_criterion = numeric(),
    rmse = numeric(),
    source = character()
  )
  
  # Loop through each row of parameters
  for(i in 1:nrow(params_df)) {
    # Extract current parameters
    current_params <- params_df[i, ]
    
    # Create model specification
    set.seed(10)
    current_spec <- decision_tree(
      tree_depth = current_params$tree_depth,
      min_n = current_params$min_n
      ) %>%
      set_engine('partykit',
      conditional_min_criterion = current_params$current_params
                 ) %>%
      set_mode('regression')
    
    # Fit model and collect metrics
    current_fit <- last_fit(current_spec, recipe, split = split)
    current_metrics <- current_fit %>% collect_metrics()
    
    # Extract RMSE value
    current_rmse <- current_metrics %>% 
      filter(.metric == "rmse") %>% 
      pull(.estimate)
    
    # Store results
    results <- results %>% add_row(
      tree_depth = current_params$tree_depth,
      min_n = current_params$min_n,
      conditional_min_criterion =current_params$conditional_min_criterion,
      rmse = current_rmse,
      source = current_params$source
    )
  }
  
  # Find the best combination
  best_params <- results %>% 
    arrange(rmse) %>% 
    slice(1)
  
  return(list(
    all_results = results,
    best_combination = best_params
  ))
}

# Example usage
results <- compare_hyperparameters(
  params_df = hyperparameters_df,
  recipe = weather_recipe,
  split = weather_split
)
print(results)
```




```{r final_spec}
final_spec <- decision_tree(tree_depth = best_rmse$tree_depth,
                            min_n = best_rmse$min_n
                            ) %>%
  # Specify the engine
  set_engine("partykit",
             conditional_min_criterion = best_rmse$conditional_min_criterion) %>%
  # Specifying mode  
  set_mode("regression")

final_spec
```

## 3. Validation  
```{r final_fit}
set.seed(10)
final_fit <- last_fit(final_spec,
                weather_recipe,
                split = weather_split)

final_fit %>%
  collect_predictions()
```
Metrics on the **test set**:
```{r}
final_fit %>%
  collect_metrics()
```
Metrics on **train set** (for curiosity and compare to test set):  
```{r}
# RMSE
final_spec %>%
  fit(grade_n ~ .,
      data = bake(weather_prep, 
                  weather_train)) %>%
  augment(new_data = bake(weather_prep, 
                          weather_train)) %>% 
  rmse(grade_n, .pred) %>%
  bind_rows(
    
    
    # R2
    final_spec %>%
      fit(grade_n ~ .,
          data = bake(weather_prep, 
                      weather_train)) %>%
      augment(new_data = bake(weather_prep, 
                              weather_train)) %>% 
      rsq(grade_n, .pred)
    
  )

```
```{r predicted vs. observed plot}
plot <- final_fit %>%
  collect_predictions() %>%
  metrica::scatter_plot(obs = grade_n,
                        pred = .pred,
                        print_eq = T,
                        print_metrics = T,
                        metrics_list = c("R2", "RMSE"),
                      # Customize metrics position
                        position_metrics = c(x=75,y=73),
                        position_eq = c(x=75, y =69))

plot
```


```{r export first plot}
ggsave(plot = plot,
       filename = "../output/ML_Grade/15_CART/M+S+O.png",
       width = 5,
       height = 5
       )
```

### Variable importance: 

```{r plotting VI}
set.seed(10)
plot_2 <- final_spec %>%
  fit(grade_n ~ .,
         data = bake(weather_prep, df)) %>%
    vip::vi() %>% 
  mutate(
    Importance = abs(Importance),
    Variable = fct_reorder(Variable, 
                           Importance)
  ) %>%
  slice_head(n = 10) %>%
  ggplot(aes(x = Importance, 
             y = Variable)) +
  geom_col() +
  scale_x_continuous(expand = c(0, 0)) +
  labs(y = NULL)

plot_2
```


```{r export second plot}
ggsave(plot = plot_2,
       filename = "../output/ML_Grade/15_CART/M+S+O_2.png",
       width = 5,
       height = 5,
       bg = "white"
       ) 
```

# 5) W + S + O
# Yield
```{r }
df <- read_sf("../data/data_complete.geojson") %>% 
  st_drop_geometry() %>% 
  as.data.frame() %>% 
  dplyr::select(planting_date, digger_date, growing_season, variety, river_basin, soil_series, row_pattern, 
                irrigation, saved_seed, 
                tillage_method=tillage_method_simplify,
                rotation = rotation_simplify_years_wo_peanut, seeding_rate, 
                grade_n, 
                alt, stage_duration_emergence:elevation,
                lat= lat.y, lon = long
                ) %>% 
  mutate_if(is.character, as.factor) %>% 
  mutate(rotation = as.factor(rotation)) %>% 
  dplyr::select(-(c(alt, rotation, seeding_rate, tillage_method))) %>% 
  dplyr::select(!(c(soil_alpha_0_5:soil_bd_30_60, 
                    soil_hb_0_5:soil_hb_30_60, 
                    soil_lambda_0_5:soil_n_30_60,
                    soil_theta_r_0_5:soil_theta_r_30_60
                    ))) %>%
  dplyr::select(-(c(SCI:NGRDI))) %>% 
  dplyr::select(grade_n, 
                #planting_date:variety, 
                #row_pattern:saved_seed, 
                #lat, lon, 
                stage_duration_emergence:stage_vp_total,
                river_basin, soil_series,
                soil_clay_0_5:soil_theta_s_30_60,
                NDVI:elevation
                ) %>% 
  #dplyr::select(-(c(NDVI:GNDVI, elevation))) %>% 
  drop_na() %>% dplyr::filter(grade_n >65) 

df
```
# ML workflow  
## 1. Pre-processing  
### a. Data split  
For data split, let's use **70% training / 30% testing**.
```{r weather_split}
# Setting seed to get reproducible results  
set.seed(27)
weather_split <- initial_split(df, prop = 0.7, 
                               strata = "grade_n")
weather_split
```
```{r weather_train}
# Setting train set 
weather_train <- training(weather_split)

weather_train
```


```{r weather_test}
# Setting test split
weather_test <- testing(weather_split)

weather_test
```
```{r data distribution}
# Combine the data frames
plot_train <- weather_train %>% mutate(Dataset = "Train")
plot_test <- weather_test %>% mutate(Dataset = "Test")
combined_data <- bind_rows(plot_train, plot_test)

# Create the density plot
ggplot(data = combined_data) +
  geom_density(aes(x = grade_n, fill = Dataset), alpha = 0.5) +
  scale_fill_manual(values = c("red", "blue")) +
  labs(x = "Yield (kg/ha)", y = "Density", fill = "Dataset")
```
  
### b. Data processing  
```{r weather_recipe}
weather_recipe <-
  # Defining predicted and predictor variables
  recipe(grade_n ~ .,
         data = weather_train) 
weather_recipe
```

```{r weather_prep}
weather_prep <- weather_recipe %>%
  prep()

weather_prep
```
## 2. Training  
### a. Model specification  
```{r hyperparameters}
cart_spec <- 
  decision_tree(tree_depth = tune(),
                min_n = tune()
                ) %>% 
  # Specify the engine
  set_engine("partykit", 
             conditional_min_criterion = tune()
             ) %>% 
  # Specifying mode  
  set_mode("regression")

cart_spec
```

### b. Hyper-parameter tuning  
  
```{r resampling_foldcv}
set.seed(15)
resampling_foldcv <- vfold_cv(weather_train, 
                              v = 5)

resampling_foldcv
```
```{r cit_param}
cit_param <- cart_spec %>% 
  extract_parameter_set_dials() %>% 
  update(conditional_min_criterion =conditional_min_criterion())

cit_param
cit_param$object
```

```{r rf_grid_result}
set.seed(45)
registerDoParallel(cores = parallel::detectCores() - 1)
rf_grid_result <- tune_race_anova(object = cart_spec,
                                   preprocessor = weather_recipe,
                                   resamples = resampling_foldcv,
                                   param_info = cit_param,
                                  grid = 50
                      )
stopImplicitCluster()
rf_grid_result$.metrics[[5]]
autoplot(rf_grid_result)
beepr::beep()
```

```{r hyperparameters selection by RMSE}
# Based on lowest RMSE
best_rmse <- rf_grid_result %>% 
  select_by_pct_loss("tree_depth",
                     metric = "rmse",
                     limit = 2
                     )%>% 
  mutate(source = "best_rmse")

best_rmse_2 <- rf_grid_result %>% 
  select_best(metric = "rmse")%>% 
  mutate(source = "best_rmse_2")

best_rmse_3 <- rf_grid_result %>% 
  select_by_one_std_err(metric = "rmse",
                        tree_depth
                        )%>% 
  mutate(source = "best_rmse_3")

```

```{r hyperparameters selection by R2}
# Based on greatest R2
best_r2 <- rf_grid_result %>% 
  select_by_pct_loss("tree_depth",
                     metric = "rsq",
                     limit = 2
                     ) %>% 
  mutate(source = "best_r2")

best_r2_2 <- rf_grid_result %>% 
  select_best(metric = "rsq")%>% 
  mutate(source = "best_r2_2")

best_r2_3 <- rf_grid_result %>% 
  select_by_one_std_err(metric = "rsq",
                        tree_depth
                        ) %>%
  mutate(source = "best_r2_3")
```
```{r comparing values}
hyperparameters_df <- best_rmse %>% 
  bind_rows(best_rmse_2, best_rmse_3)
```

```{r}
# Function to find the best model
compare_hyperparameters <- function(params_df, recipe, split) {
  # Create empty tibble to store results
  results <- tibble(
    tree_depth = numeric(),
    min_n = numeric(),
    conditional_min_criterion = numeric(),
    rmse = numeric(),
    source = character()
  )
  
  # Loop through each row of parameters
  for(i in 1:nrow(params_df)) {
    # Extract current parameters
    current_params <- params_df[i, ]
    
    # Create model specification
    set.seed(10)
    current_spec <- decision_tree(
      tree_depth = current_params$tree_depth,
      min_n = current_params$min_n
      ) %>%
      set_engine('partykit',
      conditional_min_criterion = current_params$current_params
                 ) %>%
      set_mode('regression')
    
    # Fit model and collect metrics
    current_fit <- last_fit(current_spec, recipe, split = split)
    current_metrics <- current_fit %>% collect_metrics()
    
    # Extract RMSE value
    current_rmse <- current_metrics %>% 
      filter(.metric == "rmse") %>% 
      pull(.estimate)
    
    # Store results
    results <- results %>% add_row(
      tree_depth = current_params$tree_depth,
      min_n = current_params$min_n,
      conditional_min_criterion =current_params$conditional_min_criterion,
      rmse = current_rmse,
      source = current_params$source
    )
  }
  
  # Find the best combination
  best_params <- results %>% 
    arrange(rmse) %>% 
    slice(1)
  
  return(list(
    all_results = results,
    best_combination = best_params
  ))
}

# Example usage
results <- compare_hyperparameters(
  params_df = hyperparameters_df,
  recipe = weather_recipe,
  split = weather_split
)
print(results)
```




```{r final_spec}
final_spec <- decision_tree(tree_depth = best_rmse$tree_depth,
                            min_n = best_rmse$min_n
                            ) %>%
  # Specify the engine
  set_engine("partykit",
             conditional_min_criterion = best_rmse$conditional_min_criterion) %>%
  # Specifying mode  
  set_mode("regression")

final_spec
```

## 3. Validation  
```{r final_fit}
set.seed(10)
final_fit <- last_fit(final_spec,
                weather_recipe,
                split = weather_split)

final_fit %>%
  collect_predictions()
```
Metrics on the **test set**:
```{r}
final_fit %>%
  collect_metrics()
```
Metrics on **train set** (for curiosity and compare to test set):  
```{r}
# RMSE
final_spec %>%
  fit(grade_n ~ .,
      data = bake(weather_prep, 
                  weather_train)) %>%
  augment(new_data = bake(weather_prep, 
                          weather_train)) %>% 
  rmse(grade_n, .pred) %>%
  bind_rows(
    
    
    # R2
    final_spec %>%
      fit(grade_n ~ .,
          data = bake(weather_prep, 
                      weather_train)) %>%
      augment(new_data = bake(weather_prep, 
                              weather_train)) %>% 
      rsq(grade_n, .pred)
    
  )

```
```{r predicted vs. observed plot}
plot <- final_fit %>%
  collect_predictions() %>%
  metrica::scatter_plot(obs = grade_n,
                        pred = .pred,
                        print_eq = T,
                        print_metrics = T,
                        metrics_list = c("R2", "RMSE"),
                      # Customize metrics position
                        position_metrics = c(x=75,y=75),
                        position_eq = c(x=75, y =70.5))

plot
```


```{r export first plot}
ggsave(plot = plot,
       filename = "../output/ML_Grade/15_CART/W+S+O.png",
       width = 5,
       height = 5
       )
```

### Variable importance: 

```{r plotting VI}
set.seed(10)
plot_2 <- final_spec %>%
  fit(grade_n ~ .,
         data = bake(weather_prep, df)) %>%
    vip::vi() %>% 
  mutate(
    Importance = abs(Importance),
    Variable = fct_reorder(Variable, 
                           Importance)
  ) %>%
  slice_head(n = 10) %>%
  ggplot(aes(x = Importance, 
             y = Variable)) +
  geom_col() +
  scale_x_continuous(expand = c(0, 0)) +
  labs(y = NULL)

plot_2
```


```{r export second plot}
ggsave(plot = plot_2,
       filename = "../output/ML_Grade/15_CART/W+S+O_2.png",
       width = 5,
       height = 5,
       bg = "white"
       ) 
```

# 6) M + W
# Yield
```{r }
df <- read_sf("../data/data_complete.geojson") %>% 
  st_drop_geometry() %>% 
  as.data.frame() %>% 
  dplyr::select(planting_date, digger_date, growing_season, variety, river_basin, soil_series, row_pattern, 
                irrigation, saved_seed, 
                tillage_method=tillage_method_simplify,
                rotation = rotation_simplify_years_wo_peanut, seeding_rate, 
                grade_n, 
                alt, stage_duration_emergence:elevation,
                lat= lat.y, lon = long
                ) %>% 
  mutate_if(is.character, as.factor) %>% 
  mutate(rotation = as.factor(rotation)) %>% 
  dplyr::select(-(c(alt, rotation, seeding_rate, tillage_method))) %>% 
  dplyr::select(!(c(soil_alpha_0_5:soil_bd_30_60, 
                    soil_hb_0_5:soil_hb_30_60, 
                    soil_lambda_0_5:soil_n_30_60,
                    soil_theta_r_0_5:soil_theta_r_30_60
                    ))) %>%
  dplyr::select(-(c(SCI:NGRDI))) %>% 
  dplyr::select(grade_n, planting_date:variety, 
                row_pattern:saved_seed, 
                lat, lon, 
                stage_duration_emergence:stage_vp_total#,
                #river_basin, soil_series,
                #soil_clay_0_5:soil_theta_s_30_60,
                #NDVI:elevation
                ) %>% 
  #dplyr::select(-(c(NDVI:GNDVI, elevation))) %>% 
  drop_na() %>% dplyr::filter(grade_n >65) 

df
```
# ML workflow  
## 1. Pre-processing  
### a. Data split  
For data split, let's use **70% training / 30% testing**.
```{r weather_split}
# Setting seed to get reproducible results  
set.seed(27)
weather_split <- initial_split(df, prop = 0.7, 
                               strata = "grade_n")
weather_split
```
```{r weather_train}
# Setting train set 
weather_train <- training(weather_split)

weather_train
```


```{r weather_test}
# Setting test split
weather_test <- testing(weather_split)

weather_test
```
```{r data distribution}
# Combine the data frames
plot_train <- weather_train %>% mutate(Dataset = "Train")
plot_test <- weather_test %>% mutate(Dataset = "Test")
combined_data <- bind_rows(plot_train, plot_test)

# Create the density plot
ggplot(data = combined_data) +
  geom_density(aes(x = grade_n, fill = Dataset), alpha = 0.5) +
  scale_fill_manual(values = c("red", "blue")) +
  labs(x = "Yield (kg/ha)", y = "Density", fill = "Dataset")
```
  
### b. Data processing  
```{r weather_recipe}
weather_recipe <-
  # Defining predicted and predictor variables
  recipe(grade_n ~ .,
         data = weather_train) 
weather_recipe
```

```{r weather_prep}
weather_prep <- weather_recipe %>%
  prep()

weather_prep
```
## 2. Training  
### a. Model specification  
```{r hyperparameters}
cart_spec <- 
  decision_tree(tree_depth = tune(),
                min_n = tune()
                ) %>% 
  # Specify the engine
  set_engine("partykit", 
             conditional_min_criterion = tune()
             ) %>% 
  # Specifying mode  
  set_mode("regression")

cart_spec
```

### b. Hyper-parameter tuning  
  
```{r resampling_foldcv}
set.seed(15)
resampling_foldcv <- vfold_cv(weather_train, 
                              v = 5)

resampling_foldcv
```
```{r cit_param}
cit_param <- cart_spec %>% 
  extract_parameter_set_dials() %>% 
  update(conditional_min_criterion =conditional_min_criterion())

cit_param
cit_param$object
```

```{r rf_grid_result}
set.seed(45)
registerDoParallel(cores = parallel::detectCores() - 1)
rf_grid_result <- tune_race_anova(object = cart_spec,
                                   preprocessor = weather_recipe,
                                   resamples = resampling_foldcv,
                                   param_info = cit_param,
                                  grid = 50
                      )
stopImplicitCluster()
rf_grid_result$.metrics[[5]]
autoplot(rf_grid_result)
beepr::beep()
```

```{r hyperparameters selection by RMSE}
# Based on lowest RMSE
best_rmse <- rf_grid_result %>% 
  select_by_pct_loss("tree_depth",
                     metric = "rmse",
                     limit = 2
                     )%>% 
  mutate(source = "best_rmse")

best_rmse_2 <- rf_grid_result %>% 
  select_best(metric = "rmse")%>% 
  mutate(source = "best_rmse_2")

best_rmse_3 <- rf_grid_result %>% 
  select_by_one_std_err(metric = "rmse",
                        tree_depth
                        )%>% 
  mutate(source = "best_rmse_3")

```

```{r hyperparameters selection by R2}
# Based on greatest R2
best_r2 <- rf_grid_result %>% 
  select_by_pct_loss("tree_depth",
                     metric = "rsq",
                     limit = 2
                     ) %>% 
  mutate(source = "best_r2")

best_r2_2 <- rf_grid_result %>% 
  select_best(metric = "rsq")%>% 
  mutate(source = "best_r2_2")

best_r2_3 <- rf_grid_result %>% 
  select_by_one_std_err(metric = "rsq",
                        tree_depth
                        ) %>%
  mutate(source = "best_r2_3")
```
```{r comparing values}
hyperparameters_df <- best_rmse %>% 
  bind_rows(best_rmse_2, best_rmse_3, best_r2, best_r2_2, best_r2_3)
```

```{r}
# Function to find the best model
compare_hyperparameters <- function(params_df, recipe, split) {
  # Create empty tibble to store results
  results <- tibble(
    tree_depth = numeric(),
    min_n = numeric(),
    conditional_min_criterion = numeric(),
    rmse = numeric(),
    source = character()
  )
  
  # Loop through each row of parameters
  for(i in 1:nrow(params_df)) {
    # Extract current parameters
    current_params <- params_df[i, ]
    
    # Create model specification
    set.seed(10)
    current_spec <- decision_tree(
      tree_depth = current_params$tree_depth,
      min_n = current_params$min_n
      ) %>%
      set_engine('partykit',
      conditional_min_criterion = current_params$current_params
                 ) %>%
      set_mode('regression')
    
    # Fit model and collect metrics
    current_fit <- last_fit(current_spec, recipe, split = split)
    current_metrics <- current_fit %>% collect_metrics()
    
    # Extract RMSE value
    current_rmse <- current_metrics %>% 
      filter(.metric == "rmse") %>% 
      pull(.estimate)
    
    # Store results
    results <- results %>% add_row(
      tree_depth = current_params$tree_depth,
      min_n = current_params$min_n,
      conditional_min_criterion =current_params$conditional_min_criterion,
      rmse = current_rmse,
      source = current_params$source
    )
  }
  
  # Find the best combination
  best_params <- results %>% 
    arrange(rmse) %>% 
    slice(1)
  
  return(list(
    all_results = results,
    best_combination = best_params
  ))
}

# Example usage
results <- compare_hyperparameters(
  params_df = hyperparameters_df,
  recipe = weather_recipe,
  split = weather_split
)
print(results)
```




```{r final_spec}
final_spec <- decision_tree(tree_depth = best_rmse$tree_depth,
                            min_n = best_rmse$min_n
                            ) %>%
  # Specify the engine
  set_engine("partykit",
             conditional_min_criterion = best_rmse$conditional_min_criterion) %>%
  # Specifying mode  
  set_mode("regression")

final_spec
```

## 3. Validation  
```{r final_fit}
set.seed(10)
final_fit <- last_fit(final_spec,
                weather_recipe,
                split = weather_split)

final_fit %>%
  collect_predictions()
```
Metrics on the **test set**:
```{r}
final_fit %>%
  collect_metrics()
```
Metrics on **train set** (for curiosity and compare to test set):  
```{r}
# RMSE
final_spec %>%
  fit(grade_n ~ .,
      data = bake(weather_prep, 
                  weather_train)) %>%
  augment(new_data = bake(weather_prep, 
                          weather_train)) %>% 
  rmse(grade_n, .pred) %>%
  bind_rows(
    
    
    # R2
    final_spec %>%
      fit(grade_n ~ .,
          data = bake(weather_prep, 
                      weather_train)) %>%
      augment(new_data = bake(weather_prep, 
                              weather_train)) %>% 
      rsq(grade_n, .pred)
    
  )

```
```{r predicted vs. observed plot}
plot <- final_fit %>%
  collect_predictions() %>%
  metrica::scatter_plot(obs = grade_n,
                        pred = .pred,
                        print_eq = T,
                        print_metrics = T,
                        metrics_list = c("R2", "RMSE"),
                      # Customize metrics position
                        position_metrics = c(x=75,y=73),
                        position_eq = c(x=75, y =70.5))

plot
```


```{r export first plot}
ggsave(plot = plot,
       filename = "../output/ML_Grade/15_CART/M+W.png",
       width = 5,
       height = 5
       )
```

### Variable importance: 

```{r plotting VI}
set.seed(10)
plot_2 <- final_spec %>%
  fit(grade_n ~ .,
         data = bake(weather_prep, df)) %>%
    vip::vi() %>% 
  mutate(
    Importance = abs(Importance),
    Variable = fct_reorder(Variable, 
                           Importance)
  ) %>%
  slice_head(n = 10) %>%
  ggplot(aes(x = Importance, 
             y = Variable)) +
  geom_col() +
  scale_x_continuous(expand = c(0, 0)) +
  labs(y = NULL)

plot_2
```


```{r export second plot}
ggsave(plot = plot_2,
       filename = "../output/ML_Grade/15_CART/M+W_2.png",
       width = 5,
       height = 5,
       bg = "white"
       ) 
```

# 7) M + S
# Yield
```{r }
df <- read_sf("../data/data_complete.geojson") %>% 
  st_drop_geometry() %>% 
  as.data.frame() %>% 
  dplyr::select(planting_date, digger_date, growing_season, variety, river_basin, soil_series, row_pattern, 
                irrigation, saved_seed, 
                tillage_method=tillage_method_simplify,
                rotation = rotation_simplify_years_wo_peanut, seeding_rate, 
                grade_n, 
                alt, stage_duration_emergence:elevation,
                lat= lat.y, lon = long
                ) %>% 
  mutate_if(is.character, as.factor) %>% 
  mutate(rotation = as.factor(rotation)) %>% 
  dplyr::select(-(c(alt, rotation, seeding_rate, tillage_method))) %>% 
  dplyr::select(!(c(soil_alpha_0_5:soil_bd_30_60, 
                    soil_hb_0_5:soil_hb_30_60, 
                    soil_lambda_0_5:soil_n_30_60,
                    soil_theta_r_0_5:soil_theta_r_30_60
                    ))) %>%
  dplyr::select(-(c(SCI:NGRDI))) %>% 
  dplyr::select(grade_n, planting_date:variety, 
                row_pattern:saved_seed, 
                lat, lon, 
                #stage_duration_emergence:stage_vp_total,
                river_basin, soil_series,
                soil_clay_0_5:soil_theta_s_30_60#,
                #NDVI:elevation
                ) %>% 
  #dplyr::select(-(c(NDVI:GNDVI, elevation))) %>% 
  drop_na() %>% dplyr::filter(grade_n >65) 

df
```
# ML workflow  
## 1. Pre-processing  
### a. Data split  
For data split, let's use **70% training / 30% testing**.
```{r weather_split}
# Setting seed to get reproducible results  
set.seed(27)
weather_split <- initial_split(df, prop = 0.7, 
                               strata = "grade_n")
weather_split
```
```{r weather_train}
# Setting train set 
weather_train <- training(weather_split)

weather_train
```


```{r weather_test}
# Setting test split
weather_test <- testing(weather_split)

weather_test
```
```{r data distribution}
# Combine the data frames
plot_train <- weather_train %>% mutate(Dataset = "Train")
plot_test <- weather_test %>% mutate(Dataset = "Test")
combined_data <- bind_rows(plot_train, plot_test)

# Create the density plot
ggplot(data = combined_data) +
  geom_density(aes(x = grade_n, fill = Dataset), alpha = 0.5) +
  scale_fill_manual(values = c("red", "blue")) +
  labs(x = "Yield (kg/ha)", y = "Density", fill = "Dataset")
```
  
### b. Data processing  
```{r weather_recipe}
weather_recipe <-
  # Defining predicted and predictor variables
  recipe(grade_n ~ .,
         data = weather_train) 
weather_recipe
```

```{r weather_prep}
weather_prep <- weather_recipe %>%
  prep()

weather_prep
```
## 2. Training  
### a. Model specification  
```{r hyperparameters}
cart_spec <- 
  decision_tree(tree_depth = tune(),
                min_n = tune()
                ) %>% 
  # Specify the engine
  set_engine("partykit", 
             conditional_min_criterion = tune()
             ) %>% 
  # Specifying mode  
  set_mode("regression")

cart_spec
```

### b. Hyper-parameter tuning  
  
```{r resampling_foldcv}
set.seed(15)
resampling_foldcv <- vfold_cv(weather_train, 
                              v = 5)

resampling_foldcv
```
```{r cit_param}
cit_param <- cart_spec %>% 
  extract_parameter_set_dials() %>% 
  update(conditional_min_criterion =conditional_min_criterion())

cit_param
cit_param$object
```

```{r rf_grid_result}
set.seed(45)
registerDoParallel(cores = parallel::detectCores() - 1)
rf_grid_result <- tune_race_anova(object = cart_spec,
                                   preprocessor = weather_recipe,
                                   resamples = resampling_foldcv,
                                   param_info = cit_param,
                                  grid = 50
                      )
stopImplicitCluster()
rf_grid_result$.metrics[[5]]
autoplot(rf_grid_result)
beepr::beep()
```

```{r hyperparameters selection by RMSE}
# Based on lowest RMSE
best_rmse <- rf_grid_result %>% 
  select_by_pct_loss("tree_depth",
                     metric = "rmse",
                     limit = 2
                     )%>% 
  mutate(source = "best_rmse")

best_rmse_2 <- rf_grid_result %>% 
  select_best(metric = "rmse")%>% 
  mutate(source = "best_rmse_2")

best_rmse_3 <- rf_grid_result %>% 
  select_by_one_std_err(metric = "rmse",
                        tree_depth
                        )%>% 
  mutate(source = "best_rmse_3")

```

```{r hyperparameters selection by R2}
# Based on greatest R2
best_r2 <- rf_grid_result %>% 
  select_by_pct_loss("tree_depth",
                     metric = "rsq",
                     limit = 2
                     ) %>% 
  mutate(source = "best_r2")

best_r2_2 <- rf_grid_result %>% 
  select_best(metric = "rsq")%>% 
  mutate(source = "best_r2_2")

best_r2_3 <- rf_grid_result %>% 
  select_by_one_std_err(metric = "rsq",
                        tree_depth
                        ) %>%
  mutate(source = "best_r2_3")
```
```{r comparing values}
hyperparameters_df <- best_rmse %>% 
  bind_rows(best_rmse_2, best_rmse_3, best_r2, best_r2_2, best_r2_3)
```

```{r}
# Function to find the best model
compare_hyperparameters <- function(params_df, recipe, split) {
  # Create empty tibble to store results
  results <- tibble(
    tree_depth = numeric(),
    min_n = numeric(),
    conditional_min_criterion = numeric(),
    rmse = numeric(),
    source = character()
  )
  
  # Loop through each row of parameters
  for(i in 1:nrow(params_df)) {
    # Extract current parameters
    current_params <- params_df[i, ]
    
    # Create model specification
    set.seed(10)
    current_spec <- decision_tree(
      tree_depth = current_params$tree_depth,
      min_n = current_params$min_n
      ) %>%
      set_engine('partykit',
      conditional_min_criterion = current_params$current_params
                 ) %>%
      set_mode('regression')
    
    # Fit model and collect metrics
    current_fit <- last_fit(current_spec, recipe, split = split)
    current_metrics <- current_fit %>% collect_metrics()
    
    # Extract RMSE value
    current_rmse <- current_metrics %>% 
      filter(.metric == "rmse") %>% 
      pull(.estimate)
    
    # Store results
    results <- results %>% add_row(
      tree_depth = current_params$tree_depth,
      min_n = current_params$min_n,
      conditional_min_criterion =current_params$conditional_min_criterion,
      rmse = current_rmse,
      source = current_params$source
    )
  }
  
  # Find the best combination
  best_params <- results %>% 
    arrange(rmse) %>% 
    slice(1)
  
  return(list(
    all_results = results,
    best_combination = best_params
  ))
}

# Example usage
results <- compare_hyperparameters(
  params_df = hyperparameters_df,
  recipe = weather_recipe,
  split = weather_split
)
print(results)
```




```{r final_spec}
final_spec <- decision_tree(tree_depth = best_rmse$tree_depth,
                            min_n = best_rmse$min_n
                            ) %>%
  # Specify the engine
  set_engine("partykit",
             conditional_min_criterion = best_rmse$conditional_min_criterion) %>%
  # Specifying mode  
  set_mode("regression")

final_spec
```

## 3. Validation  
```{r final_fit}
set.seed(10)
final_fit <- last_fit(final_spec,
                weather_recipe,
                split = weather_split)

final_fit %>%
  collect_predictions()
```
Metrics on the **test set**:
```{r}
final_fit %>%
  collect_metrics()
```
Metrics on **train set** (for curiosity and compare to test set):  
```{r}
# RMSE
final_spec %>%
  fit(grade_n ~ .,
      data = bake(weather_prep, 
                  weather_train)) %>%
  augment(new_data = bake(weather_prep, 
                          weather_train)) %>% 
  rmse(grade_n, .pred) %>%
  bind_rows(
    
    
    # R2
    final_spec %>%
      fit(grade_n ~ .,
          data = bake(weather_prep, 
                      weather_train)) %>%
      augment(new_data = bake(weather_prep, 
                              weather_train)) %>% 
      rsq(grade_n, .pred)
    
  )

```
```{r predicted vs. observed plot}
plot <- final_fit %>%
  collect_predictions() %>%
  metrica::scatter_plot(obs = grade_n,
                        pred = .pred,
                        print_eq = T,
                        print_metrics = T,
                        metrics_list = c("R2", "RMSE"),
                      # Customize metrics position
                        position_metrics = c(x=75,y=73),
                        position_eq = c(x=75, y =70.5))

plot
```


```{r export first plot}
ggsave(plot = plot,
       filename = "../output/ML_Grade/15_CART/M+S.png",
       width = 5,
       height = 5
       )
```

### Variable importance: 

```{r plotting VI}
set.seed(10)
plot_2 <- final_spec %>%
  fit(grade_n ~ .,
         data = bake(weather_prep, df)) %>%
    vip::vi() %>% 
  mutate(
    Importance = abs(Importance),
    Variable = fct_reorder(Variable, 
                           Importance)
  ) %>%
  slice_head(n = 10) %>%
  ggplot(aes(x = Importance, 
             y = Variable)) +
  geom_col() +
  scale_x_continuous(expand = c(0, 0)) +
  labs(y = NULL)

plot_2
```


```{r export second plot}
ggsave(plot = plot_2,
       filename = "../output/ML_Grade/15_CART/M+S_2.png",
       width = 5,
       height = 5,
       bg = "white"
       ) 
```

# 8) M + O
# Yield
```{r }
df <- read_sf("../data/data_complete.geojson") %>% 
  st_drop_geometry() %>% 
  as.data.frame() %>% 
  dplyr::select(planting_date, digger_date, growing_season, variety, river_basin, soil_series, row_pattern, 
                irrigation, saved_seed, 
                tillage_method=tillage_method_simplify,
                rotation = rotation_simplify_years_wo_peanut, seeding_rate, 
                grade_n, 
                alt, stage_duration_emergence:elevation,
                lat= lat.y, lon = long
                ) %>% 
  mutate_if(is.character, as.factor) %>% 
  mutate(rotation = as.factor(rotation)) %>% 
  dplyr::select(-(c(alt, rotation, seeding_rate, tillage_method))) %>% 
  dplyr::select(!(c(soil_alpha_0_5:soil_bd_30_60, 
                    soil_hb_0_5:soil_hb_30_60, 
                    soil_lambda_0_5:soil_n_30_60,
                    soil_theta_r_0_5:soil_theta_r_30_60
                    ))) %>%
  dplyr::select(-(c(SCI:NGRDI))) %>% 
  dplyr::select(grade_n, planting_date:variety, 
                row_pattern:saved_seed, 
                lat, lon, 
                #stage_duration_emergence:stage_vp_total,
                #river_basin, soil_series,
                #soil_clay_0_5:soil_theta_s_30_60,
                NDVI:elevation
                ) %>% 
  #dplyr::select(-(c(NDVI:GNDVI, elevation))) %>% 
  drop_na() %>% dplyr::filter(grade_n >65) 

df
```
# ML workflow  
## 1. Pre-processing  
### a. Data split  
For data split, let's use **70% training / 30% testing**.
```{r weather_split}
# Setting seed to get reproducible results  
set.seed(27)
weather_split <- initial_split(df, prop = 0.7, 
                               strata = "grade_n")
weather_split
```
```{r weather_train}
# Setting train set 
weather_train <- training(weather_split)

weather_train
```


```{r weather_test}
# Setting test split
weather_test <- testing(weather_split)

weather_test
```
```{r data distribution}
# Combine the data frames
plot_train <- weather_train %>% mutate(Dataset = "Train")
plot_test <- weather_test %>% mutate(Dataset = "Test")
combined_data <- bind_rows(plot_train, plot_test)

# Create the density plot
ggplot(data = combined_data) +
  geom_density(aes(x = grade_n, fill = Dataset), alpha = 0.5) +
  scale_fill_manual(values = c("red", "blue")) +
  labs(x = "Yield (kg/ha)", y = "Density", fill = "Dataset")
```
  
### b. Data processing  
```{r weather_recipe}
weather_recipe <-
  # Defining predicted and predictor variables
  recipe(grade_n ~ .,
         data = weather_train) 
weather_recipe
```

```{r weather_prep}
weather_prep <- weather_recipe %>%
  prep()

weather_prep
```
## 2. Training  
### a. Model specification  
```{r hyperparameters}
cart_spec <- 
  decision_tree(tree_depth = tune(),
                min_n = tune()
                ) %>% 
  # Specify the engine
  set_engine("partykit", 
             conditional_min_criterion = tune()
             ) %>% 
  # Specifying mode  
  set_mode("regression")

cart_spec
```

### b. Hyper-parameter tuning  
  
```{r resampling_foldcv}
set.seed(15)
resampling_foldcv <- vfold_cv(weather_train, 
                              v = 5)

resampling_foldcv
```
```{r cit_param}
cit_param <- cart_spec %>% 
  extract_parameter_set_dials() %>% 
  update(conditional_min_criterion =conditional_min_criterion())

cit_param
cit_param$object
```

```{r rf_grid_result}
set.seed(45)
registerDoParallel(cores = parallel::detectCores() - 1)
rf_grid_result <- tune_race_anova(object = cart_spec,
                                   preprocessor = weather_recipe,
                                   resamples = resampling_foldcv,
                                   param_info = cit_param,
                                  grid = 50
                      )
stopImplicitCluster()
rf_grid_result$.metrics[[5]]
autoplot(rf_grid_result)
beepr::beep()
```

```{r hyperparameters selection by RMSE}
# Based on lowest RMSE
best_rmse <- rf_grid_result %>% 
  select_by_pct_loss("tree_depth",
                     metric = "rmse",
                     limit = 2
                     )%>% 
  mutate(source = "best_rmse")

best_rmse_2 <- rf_grid_result %>% 
  select_best(metric = "rmse")%>% 
  mutate(source = "best_rmse_2")

best_rmse_3 <- rf_grid_result %>% 
  select_by_one_std_err(metric = "rmse",
                        tree_depth
                        )%>% 
  mutate(source = "best_rmse_3")

```

```{r hyperparameters selection by R2}
# Based on greatest R2
best_r2 <- rf_grid_result %>% 
  select_by_pct_loss("tree_depth",
                     metric = "rsq",
                     limit = 2
                     ) %>% 
  mutate(source = "best_r2")

best_r2_2 <- rf_grid_result %>% 
  select_best(metric = "rsq")%>% 
  mutate(source = "best_r2_2")

best_r2_3 <- rf_grid_result %>% 
  select_by_one_std_err(metric = "rsq",
                        tree_depth
                        ) %>%
  mutate(source = "best_r2_3")
```
```{r comparing values}
hyperparameters_df <- best_rmse %>% 
  bind_rows(best_rmse_2, best_rmse_3, best_r2, best_r2_2, best_r2_3)
```

```{r}
# Function to find the best model
compare_hyperparameters <- function(params_df, recipe, split) {
  # Create empty tibble to store results
  results <- tibble(
    tree_depth = numeric(),
    min_n = numeric(),
    conditional_min_criterion = numeric(),
    rmse = numeric(),
    source = character()
  )
  
  # Loop through each row of parameters
  for(i in 1:nrow(params_df)) {
    # Extract current parameters
    current_params <- params_df[i, ]
    
    # Create model specification
    set.seed(10)
    current_spec <- decision_tree(
      tree_depth = current_params$tree_depth,
      min_n = current_params$min_n
      ) %>%
      set_engine('partykit',
      conditional_min_criterion = current_params$current_params
                 ) %>%
      set_mode('regression')
    
    # Fit model and collect metrics
    current_fit <- last_fit(current_spec, recipe, split = split)
    current_metrics <- current_fit %>% collect_metrics()
    
    # Extract RMSE value
    current_rmse <- current_metrics %>% 
      filter(.metric == "rmse") %>% 
      pull(.estimate)
    
    # Store results
    results <- results %>% add_row(
      tree_depth = current_params$tree_depth,
      min_n = current_params$min_n,
      conditional_min_criterion =current_params$conditional_min_criterion,
      rmse = current_rmse,
      source = current_params$source
    )
  }
  
  # Find the best combination
  best_params <- results %>% 
    arrange(rmse) %>% 
    slice(1)
  
  return(list(
    all_results = results,
    best_combination = best_params
  ))
}

# Example usage
results <- compare_hyperparameters(
  params_df = hyperparameters_df,
  recipe = weather_recipe,
  split = weather_split
)
print(results)
```




```{r final_spec}
final_spec <- decision_tree(tree_depth = best_rmse$tree_depth,
                            min_n = best_rmse$min_n
                            ) %>%
  # Specify the engine
  set_engine("partykit",
             conditional_min_criterion = best_rmse$conditional_min_criterion) %>%
  # Specifying mode  
  set_mode("regression")

final_spec
```

## 3. Validation  
```{r final_fit}
set.seed(10)
final_fit <- last_fit(final_spec,
                weather_recipe,
                split = weather_split)

final_fit %>%
  collect_predictions()
```
Metrics on the **test set**:
```{r}
final_fit %>%
  collect_metrics()
```
Metrics on **train set** (for curiosity and compare to test set):  
```{r}
# RMSE
final_spec %>%
  fit(grade_n ~ .,
      data = bake(weather_prep, 
                  weather_train)) %>%
  augment(new_data = bake(weather_prep, 
                          weather_train)) %>% 
  rmse(grade_n, .pred) %>%
  bind_rows(
    
    
    # R2
    final_spec %>%
      fit(grade_n ~ .,
          data = bake(weather_prep, 
                      weather_train)) %>%
      augment(new_data = bake(weather_prep, 
                              weather_train)) %>% 
      rsq(grade_n, .pred)
    
  )

```
```{r predicted vs. observed plot}
plot <- final_fit %>%
  collect_predictions() %>%
  metrica::scatter_plot(obs = grade_n,
                        pred = .pred,
                        print_eq = T,
                        print_metrics = T,
                        metrics_list = c("R2", "RMSE"),
                      # Customize metrics position
                        position_metrics = c(x=75,y=73),
                        position_eq = c(x=75, y =70.8))

plot
```


```{r export first plot}
ggsave(plot = plot,
       filename = "../output/ML_Grade/15_CART/M+O.png",
       width = 5,
       height = 5
       )
```

### Variable importance: 

```{r plotting VI}
set.seed(10)
plot_2 <- final_spec %>%
  fit(grade_n ~ .,
         data = bake(weather_prep, df)) %>%
    vip::vi() %>% 
  mutate(
    Importance = abs(Importance),
    Variable = fct_reorder(Variable, 
                           Importance)
  ) %>%
  slice_head(n = 10) %>%
  ggplot(aes(x = Importance, 
             y = Variable)) +
  geom_col() +
  scale_x_continuous(expand = c(0, 0)) +
  labs(y = NULL)

plot_2
```


```{r export second plot}
ggsave(plot = plot_2,
       filename = "../output/ML_Grade/15_CART/M+O_2.png",
       width = 5,
       height = 5,
       bg = "white"
       ) 
```

# 9) W + S
# Yield
```{r }
df <- read_sf("../data/data_complete.geojson") %>% 
  st_drop_geometry() %>% 
  as.data.frame() %>% 
  dplyr::select(planting_date, digger_date, growing_season, variety, river_basin, soil_series, row_pattern, 
                irrigation, saved_seed, 
                tillage_method=tillage_method_simplify,
                rotation = rotation_simplify_years_wo_peanut, seeding_rate, 
                grade_n, 
                alt, stage_duration_emergence:elevation,
                lat= lat.y, lon = long
                ) %>% 
  mutate_if(is.character, as.factor) %>% 
  mutate(rotation = as.factor(rotation)) %>% 
  dplyr::select(-(c(alt, rotation, seeding_rate, tillage_method))) %>% 
  dplyr::select(!(c(soil_alpha_0_5:soil_bd_30_60, 
                    soil_hb_0_5:soil_hb_30_60, 
                    soil_lambda_0_5:soil_n_30_60,
                    soil_theta_r_0_5:soil_theta_r_30_60
                    ))) %>%
  dplyr::select(-(c(SCI:NGRDI))) %>% 
  dplyr::select(grade_n, 
                #planting_date:variety, 
                #row_pattern:saved_seed, 
                #lat, lon, 
                stage_duration_emergence:stage_vp_total,
                river_basin, soil_series,
                soil_clay_0_5:soil_theta_s_30_60#,
                #NDVI:elevation
                ) %>% 
  #dplyr::select(-(c(NDVI:GNDVI, elevation))) %>% 
  drop_na() %>% dplyr::filter(grade_n >65) 

df
```
# ML workflow  
## 1. Pre-processing  
### a. Data split  
For data split, let's use **70% training / 30% testing**.
```{r weather_split}
# Setting seed to get reproducible results  
set.seed(27)
weather_split <- initial_split(df, prop = 0.7, 
                               strata = "grade_n")
weather_split
```
```{r weather_train}
# Setting train set 
weather_train <- training(weather_split)

weather_train
```


```{r weather_test}
# Setting test split
weather_test <- testing(weather_split)

weather_test
```
```{r data distribution}
# Combine the data frames
plot_train <- weather_train %>% mutate(Dataset = "Train")
plot_test <- weather_test %>% mutate(Dataset = "Test")
combined_data <- bind_rows(plot_train, plot_test)

# Create the density plot
ggplot(data = combined_data) +
  geom_density(aes(x = grade_n, fill = Dataset), alpha = 0.5) +
  scale_fill_manual(values = c("red", "blue")) +
  labs(x = "Yield (kg/ha)", y = "Density", fill = "Dataset")
```
  
### b. Data processing  
```{r weather_recipe}
weather_recipe <-
  # Defining predicted and predictor variables
  recipe(grade_n ~ .,
         data = weather_train) 
weather_recipe
```

```{r weather_prep}
weather_prep <- weather_recipe %>%
  prep()

weather_prep
```
## 2. Training  
### a. Model specification  
```{r hyperparameters}
cart_spec <- 
  decision_tree(tree_depth = tune(),
                min_n = tune()
                ) %>% 
  # Specify the engine
  set_engine("partykit", 
             conditional_min_criterion = tune()
             ) %>% 
  # Specifying mode  
  set_mode("regression")

cart_spec
```

### b. Hyper-parameter tuning  
  
```{r resampling_foldcv}
set.seed(15)
resampling_foldcv <- vfold_cv(weather_train, 
                              v = 5)

resampling_foldcv
```
```{r cit_param}
cit_param <- cart_spec %>% 
  extract_parameter_set_dials() %>% 
  update(conditional_min_criterion =conditional_min_criterion())

cit_param
cit_param$object
```

```{r rf_grid_result}
set.seed(45)
registerDoParallel(cores = parallel::detectCores() - 1)
rf_grid_result <- tune_race_anova(object = cart_spec,
                                   preprocessor = weather_recipe,
                                   resamples = resampling_foldcv,
                                   param_info = cit_param,
                                  grid = 50
                      )
stopImplicitCluster()
rf_grid_result$.metrics[[5]]
autoplot(rf_grid_result)
beepr::beep()
```

```{r hyperparameters selection by RMSE}
# Based on lowest RMSE
best_rmse <- rf_grid_result %>% 
  select_by_pct_loss("tree_depth",
                     metric = "rmse",
                     limit = 2
                     )%>% 
  mutate(source = "best_rmse")

best_rmse_2 <- rf_grid_result %>% 
  select_best(metric = "rmse")%>% 
  mutate(source = "best_rmse_2")

best_rmse_3 <- rf_grid_result %>% 
  select_by_one_std_err(metric = "rmse",
                        tree_depth
                        )%>% 
  mutate(source = "best_rmse_3")

```

```{r hyperparameters selection by R2}
# Based on greatest R2
best_r2 <- rf_grid_result %>% 
  select_by_pct_loss("tree_depth",
                     metric = "rsq",
                     limit = 2
                     ) %>% 
  mutate(source = "best_r2")

best_r2_2 <- rf_grid_result %>% 
  select_best(metric = "rsq")%>% 
  mutate(source = "best_r2_2")

best_r2_3 <- rf_grid_result %>% 
  select_by_one_std_err(metric = "rsq",
                        tree_depth
                        ) %>%
  mutate(source = "best_r2_3")
```
```{r comparing values}
hyperparameters_df <- best_rmse %>% 
  bind_rows(best_rmse_2, best_rmse_3, best_r2, best_r2_2, best_r2_3)
```

```{r}
# Function to find the best model
compare_hyperparameters <- function(params_df, recipe, split) {
  # Create empty tibble to store results
  results <- tibble(
    tree_depth = numeric(),
    min_n = numeric(),
    conditional_min_criterion = numeric(),
    rmse = numeric(),
    source = character()
  )
  
  # Loop through each row of parameters
  for(i in 1:nrow(params_df)) {
    # Extract current parameters
    current_params <- params_df[i, ]
    
    # Create model specification
    set.seed(10)
    current_spec <- decision_tree(
      tree_depth = current_params$tree_depth,
      min_n = current_params$min_n
      ) %>%
      set_engine('partykit',
      conditional_min_criterion = current_params$current_params
                 ) %>%
      set_mode('regression')
    
    # Fit model and collect metrics
    current_fit <- last_fit(current_spec, recipe, split = split)
    current_metrics <- current_fit %>% collect_metrics()
    
    # Extract RMSE value
    current_rmse <- current_metrics %>% 
      filter(.metric == "rmse") %>% 
      pull(.estimate)
    
    # Store results
    results <- results %>% add_row(
      tree_depth = current_params$tree_depth,
      min_n = current_params$min_n,
      conditional_min_criterion =current_params$conditional_min_criterion,
      rmse = current_rmse,
      source = current_params$source
    )
  }
  
  # Find the best combination
  best_params <- results %>% 
    arrange(rmse) %>% 
    slice(1)
  
  return(list(
    all_results = results,
    best_combination = best_params
  ))
}

# Example usage
results <- compare_hyperparameters(
  params_df = hyperparameters_df,
  recipe = weather_recipe,
  split = weather_split
)
print(results)
```




```{r final_spec}
final_spec <- decision_tree(tree_depth = best_rmse$tree_depth,
                            min_n = best_rmse$min_n
                            ) %>%
  # Specify the engine
  set_engine("partykit",
             conditional_min_criterion = best_rmse$conditional_min_criterion) %>%
  # Specifying mode  
  set_mode("regression")

final_spec
```

## 3. Validation  
```{r final_fit}
set.seed(10)
final_fit <- last_fit(final_spec,
                weather_recipe,
                split = weather_split)

final_fit %>%
  collect_predictions()
```
Metrics on the **test set**:
```{r}
final_fit %>%
  collect_metrics()
```
Metrics on **train set** (for curiosity and compare to test set):  
```{r}
# RMSE
final_spec %>%
  fit(grade_n ~ .,
      data = bake(weather_prep, 
                  weather_train)) %>%
  augment(new_data = bake(weather_prep, 
                          weather_train)) %>% 
  rmse(grade_n, .pred) %>%
  bind_rows(
    
    
    # R2
    final_spec %>%
      fit(grade_n ~ .,
          data = bake(weather_prep, 
                      weather_train)) %>%
      augment(new_data = bake(weather_prep, 
                              weather_train)) %>% 
      rsq(grade_n, .pred)
    
  )

```
```{r predicted vs. observed plot}
plot <- final_fit %>%
  collect_predictions() %>%
  metrica::scatter_plot(obs = grade_n,
                        pred = .pred,
                        print_eq = T,
                        print_metrics = T,
                        metrics_list = c("R2", "RMSE"),
                      # Customize metrics position
                        position_metrics = c(x=75,y=73),
                        position_eq = c(x=75, y =70.5))

plot
```


```{r export first plot}
ggsave(plot = plot,
       filename = "../output/ML_Grade/15_CART/W+S.png",
       width = 5,
       height = 5
       )
```

### Variable importance: 

```{r plotting VI}
set.seed(10)
plot_2 <- final_spec %>%
  fit(grade_n ~ .,
         data = bake(weather_prep, df)) %>%
    vip::vi() %>% 
  mutate(
    Importance = abs(Importance),
    Variable = fct_reorder(Variable, 
                           Importance)
  ) %>%
  slice_head(n = 10) %>%
  ggplot(aes(x = Importance, 
             y = Variable)) +
  geom_col() +
  scale_x_continuous(expand = c(0, 0)) +
  labs(y = NULL)

plot_2
```


```{r export second plot}
ggsave(plot = plot_2,
       filename = "../output/ML_Grade/15_CART/W+S_2.png",
       width = 5,
       height = 5,
       bg = "white"
       ) 
```

# 10) W + O
# Yield
```{r }
df <- read_sf("../data/data_complete.geojson") %>% 
  st_drop_geometry() %>% 
  as.data.frame() %>% 
  dplyr::select(planting_date, digger_date, growing_season, variety, river_basin, soil_series, row_pattern, 
                irrigation, saved_seed, 
                tillage_method=tillage_method_simplify,
                rotation = rotation_simplify_years_wo_peanut, seeding_rate, 
                grade_n, 
                alt, stage_duration_emergence:elevation,
                lat= lat.y, lon = long
                ) %>% 
  mutate_if(is.character, as.factor) %>% 
  mutate(rotation = as.factor(rotation)) %>% 
  dplyr::select(-(c(alt, rotation, seeding_rate, tillage_method))) %>% 
  dplyr::select(!(c(soil_alpha_0_5:soil_bd_30_60, 
                    soil_hb_0_5:soil_hb_30_60, 
                    soil_lambda_0_5:soil_n_30_60,
                    soil_theta_r_0_5:soil_theta_r_30_60
                    ))) %>%
  dplyr::select(-(c(SCI:NGRDI))) %>% 
  dplyr::select(grade_n, 
                #planting_date:variety, 
                #row_pattern:saved_seed, 
                #lat, lon, 
                stage_duration_emergence:stage_vp_total,
                #river_basin, soil_series,
                #soil_clay_0_5:soil_theta_s_30_60,
                NDVI:elevation
                ) %>% 
  #dplyr::select(-(c(NDVI:GNDVI, elevation))) %>% 
  drop_na() %>% dplyr::filter(grade_n >65) 

df
```
# ML workflow  
## 1. Pre-processing  
### a. Data split  
For data split, let's use **70% training / 30% testing**.
```{r weather_split}
# Setting seed to get reproducible results  
set.seed(27)
weather_split <- initial_split(df, prop = 0.7, 
                               strata = "grade_n")
weather_split
```
```{r weather_train}
# Setting train set 
weather_train <- training(weather_split)

weather_train
```


```{r weather_test}
# Setting test split
weather_test <- testing(weather_split)

weather_test
```
```{r data distribution}
# Combine the data frames
plot_train <- weather_train %>% mutate(Dataset = "Train")
plot_test <- weather_test %>% mutate(Dataset = "Test")
combined_data <- bind_rows(plot_train, plot_test)

# Create the density plot
ggplot(data = combined_data) +
  geom_density(aes(x = grade_n, fill = Dataset), alpha = 0.5) +
  scale_fill_manual(values = c("red", "blue")) +
  labs(x = "Yield (kg/ha)", y = "Density", fill = "Dataset")
```
  
### b. Data processing  
```{r weather_recipe}
weather_recipe <-
  # Defining predicted and predictor variables
  recipe(grade_n ~ .,
         data = weather_train) 
weather_recipe
```

```{r weather_prep}
weather_prep <- weather_recipe %>%
  prep()

weather_prep
```
## 2. Training  
### a. Model specification  
```{r hyperparameters}
cart_spec <- 
  decision_tree(tree_depth = tune(),
                min_n = tune()
                ) %>% 
  # Specify the engine
  set_engine("partykit", 
             conditional_min_criterion = tune()
             ) %>% 
  # Specifying mode  
  set_mode("regression")

cart_spec
```

### b. Hyper-parameter tuning  
  
```{r resampling_foldcv}
set.seed(15)
resampling_foldcv <- vfold_cv(weather_train, 
                              v = 5)

resampling_foldcv
```
```{r cit_param}
cit_param <- cart_spec %>% 
  extract_parameter_set_dials() %>% 
  update(conditional_min_criterion =conditional_min_criterion())

cit_param
cit_param$object
```

```{r rf_grid_result}
set.seed(45)
registerDoParallel(cores = parallel::detectCores() - 1)
rf_grid_result <- tune_race_anova(object = cart_spec,
                                   preprocessor = weather_recipe,
                                   resamples = resampling_foldcv,
                                   param_info = cit_param,
                                  grid = 50
                      )
stopImplicitCluster()
rf_grid_result$.metrics[[5]]
autoplot(rf_grid_result)
beepr::beep()
```

```{r hyperparameters selection by RMSE}
# Based on lowest RMSE
best_rmse <- rf_grid_result %>% 
  select_by_pct_loss("tree_depth",
                     metric = "rmse",
                     limit = 2
                     )%>% 
  mutate(source = "best_rmse")

best_rmse_2 <- rf_grid_result %>% 
  select_best(metric = "rmse")%>% 
  mutate(source = "best_rmse_2")

best_rmse_3 <- rf_grid_result %>% 
  select_by_one_std_err(metric = "rmse",
                        tree_depth
                        )%>% 
  mutate(source = "best_rmse_3")

```

```{r hyperparameters selection by R2}
# Based on greatest R2
best_r2 <- rf_grid_result %>% 
  select_by_pct_loss("tree_depth",
                     metric = "rsq",
                     limit = 2
                     ) %>% 
  mutate(source = "best_r2")

best_r2_2 <- rf_grid_result %>% 
  select_best(metric = "rsq")%>% 
  mutate(source = "best_r2_2")

best_r2_3 <- rf_grid_result %>% 
  select_by_one_std_err(metric = "rsq",
                        tree_depth
                        ) %>%
  mutate(source = "best_r2_3")
```
```{r comparing values}
hyperparameters_df <- best_rmse %>% 
  bind_rows(best_rmse_2, best_rmse_3, best_r2, best_r2_2, best_r2_3)
```

```{r}
# Function to find the best model
compare_hyperparameters <- function(params_df, recipe, split) {
  # Create empty tibble to store results
  results <- tibble(
    tree_depth = numeric(),
    min_n = numeric(),
    conditional_min_criterion = numeric(),
    rmse = numeric(),
    source = character()
  )
  
  # Loop through each row of parameters
  for(i in 1:nrow(params_df)) {
    # Extract current parameters
    current_params <- params_df[i, ]
    
    # Create model specification
    set.seed(10)
    current_spec <- decision_tree(
      tree_depth = current_params$tree_depth,
      min_n = current_params$min_n
      ) %>%
      set_engine('partykit',
      conditional_min_criterion = current_params$current_params
                 ) %>%
      set_mode('regression')
    
    # Fit model and collect metrics
    current_fit <- last_fit(current_spec, recipe, split = split)
    current_metrics <- current_fit %>% collect_metrics()
    
    # Extract RMSE value
    current_rmse <- current_metrics %>% 
      filter(.metric == "rmse") %>% 
      pull(.estimate)
    
    # Store results
    results <- results %>% add_row(
      tree_depth = current_params$tree_depth,
      min_n = current_params$min_n,
      conditional_min_criterion =current_params$conditional_min_criterion,
      rmse = current_rmse,
      source = current_params$source
    )
  }
  
  # Find the best combination
  best_params <- results %>% 
    arrange(rmse) %>% 
    slice(1)
  
  return(list(
    all_results = results,
    best_combination = best_params
  ))
}

# Example usage
results <- compare_hyperparameters(
  params_df = hyperparameters_df,
  recipe = weather_recipe,
  split = weather_split
)
print(results)
```




```{r final_spec}
final_spec <- decision_tree(tree_depth = best_rmse$tree_depth,
                            min_n = best_rmse$min_n
                            ) %>%
  # Specify the engine
  set_engine("partykit",
             conditional_min_criterion = best_rmse$conditional_min_criterion) %>%
  # Specifying mode  
  set_mode("regression")

final_spec
```

## 3. Validation  
```{r final_fit}
set.seed(10)
final_fit <- last_fit(final_spec,
                weather_recipe,
                split = weather_split)

final_fit %>%
  collect_predictions()
```
Metrics on the **test set**:
```{r}
final_fit %>%
  collect_metrics()
```
Metrics on **train set** (for curiosity and compare to test set):  
```{r}
# RMSE
final_spec %>%
  fit(grade_n ~ .,
      data = bake(weather_prep, 
                  weather_train)) %>%
  augment(new_data = bake(weather_prep, 
                          weather_train)) %>% 
  rmse(grade_n, .pred) %>%
  bind_rows(
    
    
    # R2
    final_spec %>%
      fit(grade_n ~ .,
          data = bake(weather_prep, 
                      weather_train)) %>%
      augment(new_data = bake(weather_prep, 
                              weather_train)) %>% 
      rsq(grade_n, .pred)
    
  )

```
```{r predicted vs. observed plot}
plot <- final_fit %>%
  collect_predictions() %>%
  metrica::scatter_plot(obs = grade_n,
                        pred = .pred,
                        print_eq = T,
                        print_metrics = T,
                        metrics_list = c("R2", "RMSE"),
                      # Customize metrics position
                        position_metrics = c(x=75,y=73),
                        position_eq = c(x=75, y =71))

plot
```


```{r export first plot}
ggsave(plot = plot,
       filename = "../output/ML_Grade/15_CART/W+O.png",
       width = 5,
       height = 5
       )
```

### Variable importance: 

```{r plotting VI}
set.seed(10)
plot_2 <- final_spec %>%
  fit(grade_n ~ .,
         data = bake(weather_prep, df)) %>%
    vip::vi() %>% 
  mutate(
    Importance = abs(Importance),
    Variable = fct_reorder(Variable, 
                           Importance)
  ) %>%
  slice_head(n = 10) %>%
  ggplot(aes(x = Importance, 
             y = Variable)) +
  geom_col() +
  scale_x_continuous(expand = c(0, 0)) +
  labs(y = NULL)

plot_2
```


```{r export second plot}
ggsave(plot = plot_2,
       filename = "../output/ML_Grade/15_CART/W+O_2.png",
       width = 5,
       height = 5,
       bg = "white"
       ) 
```

# 11) S + O
# Yield
```{r }
df <- read_sf("../data/data_complete.geojson") %>% 
  st_drop_geometry() %>% 
  as.data.frame() %>% 
  dplyr::select(planting_date, digger_date, growing_season, variety, river_basin, soil_series, row_pattern, 
                irrigation, saved_seed, 
                tillage_method=tillage_method_simplify,
                rotation = rotation_simplify_years_wo_peanut, seeding_rate, 
                grade_n, 
                alt, stage_duration_emergence:elevation,
                lat= lat.y, lon = long
                ) %>% 
  mutate_if(is.character, as.factor) %>% 
  mutate(rotation = as.factor(rotation)) %>% 
  dplyr::select(-(c(alt, rotation, seeding_rate, tillage_method))) %>% 
  dplyr::select(!(c(soil_alpha_0_5:soil_bd_30_60, 
                    soil_hb_0_5:soil_hb_30_60, 
                    soil_lambda_0_5:soil_n_30_60,
                    soil_theta_r_0_5:soil_theta_r_30_60
                    ))) %>%
  dplyr::select(-(c(SCI:NGRDI))) %>% 
  dplyr::select(grade_n, 
                #planting_date:variety, 
                #row_pattern:saved_seed, 
                #lat, lon, 
                #stage_duration_emergence:stage_vp_total,
                river_basin, soil_series,
                soil_clay_0_5:soil_theta_s_30_60,
                NDVI:elevation
                ) %>% 
  #dplyr::select(-(c(NDVI:GNDVI, elevation))) %>% 
  drop_na() %>% dplyr::filter(grade_n >65) 

df
```
# ML workflow  
## 1. Pre-processing  
### a. Data split  
For data split, let's use **70% training / 30% testing**.
```{r weather_split}
# Setting seed to get reproducible results  
set.seed(27)
weather_split <- initial_split(df, prop = 0.7, 
                               strata = "grade_n")
weather_split
```
```{r weather_train}
# Setting train set 
weather_train <- training(weather_split)

weather_train
```


```{r weather_test}
# Setting test split
weather_test <- testing(weather_split)

weather_test
```
```{r data distribution}
# Combine the data frames
plot_train <- weather_train %>% mutate(Dataset = "Train")
plot_test <- weather_test %>% mutate(Dataset = "Test")
combined_data <- bind_rows(plot_train, plot_test)

# Create the density plot
ggplot(data = combined_data) +
  geom_density(aes(x = grade_n, fill = Dataset), alpha = 0.5) +
  scale_fill_manual(values = c("red", "blue")) +
  labs(x = "Yield (kg/ha)", y = "Density", fill = "Dataset")
```
  
### b. Data processing  
```{r weather_recipe}
weather_recipe <-
  # Defining predicted and predictor variables
  recipe(grade_n ~ .,
         data = weather_train) 
weather_recipe
```

```{r weather_prep}
weather_prep <- weather_recipe %>%
  prep()

weather_prep
```
## 2. Training  
### a. Model specification  
```{r hyperparameters}
cart_spec <- 
  decision_tree(tree_depth = tune(),
                min_n = tune()
                ) %>% 
  # Specify the engine
  set_engine("partykit", 
             conditional_min_criterion = tune()
             ) %>% 
  # Specifying mode  
  set_mode("regression")

cart_spec
```

### b. Hyper-parameter tuning  
  
```{r resampling_foldcv}
set.seed(15)
resampling_foldcv <- vfold_cv(weather_train, 
                              v = 5)

resampling_foldcv
```
```{r cit_param}
cit_param <- cart_spec %>% 
  extract_parameter_set_dials() %>% 
  update(conditional_min_criterion =conditional_min_criterion())

cit_param
cit_param$object
```

```{r rf_grid_result}
set.seed(45)
registerDoParallel(cores = parallel::detectCores() - 1)
rf_grid_result <- tune_race_anova(object = cart_spec,
                                   preprocessor = weather_recipe,
                                   resamples = resampling_foldcv,
                                   param_info = cit_param,
                                  grid = 50
                      )
stopImplicitCluster()
rf_grid_result$.metrics[[5]]
autoplot(rf_grid_result)
beepr::beep()
```

```{r hyperparameters selection by RMSE}
# Based on lowest RMSE
best_rmse <- rf_grid_result %>% 
  select_by_pct_loss("tree_depth",
                     metric = "rmse",
                     limit = 2
                     )%>% 
  mutate(source = "best_rmse")

best_rmse_2 <- rf_grid_result %>% 
  select_best(metric = "rmse")%>% 
  mutate(source = "best_rmse_2")

best_rmse_3 <- rf_grid_result %>% 
  select_by_one_std_err(metric = "rmse",
                        tree_depth
                        )%>% 
  mutate(source = "best_rmse_3")

```

```{r hyperparameters selection by R2}
# Based on greatest R2
best_r2 <- rf_grid_result %>% 
  select_by_pct_loss("tree_depth",
                     metric = "rsq",
                     limit = 2
                     ) %>% 
  mutate(source = "best_r2")

best_r2_2 <- rf_grid_result %>% 
  select_best(metric = "rsq")%>% 
  mutate(source = "best_r2_2")

best_r2_3 <- rf_grid_result %>% 
  select_by_one_std_err(metric = "rsq",
                        tree_depth
                        ) %>%
  mutate(source = "best_r2_3")
```
```{r comparing values}
hyperparameters_df <- best_rmse %>% 
  bind_rows(best_rmse_2, best_rmse_3, best_r2, best_r2_2, best_r2_3)
```

```{r}
# Function to find the best model
compare_hyperparameters <- function(params_df, recipe, split) {
  # Create empty tibble to store results
  results <- tibble(
    tree_depth = numeric(),
    min_n = numeric(),
    conditional_min_criterion = numeric(),
    rmse = numeric(),
    source = character()
  )
  
  # Loop through each row of parameters
  for(i in 1:nrow(params_df)) {
    # Extract current parameters
    current_params <- params_df[i, ]
    
    # Create model specification
    set.seed(10)
    current_spec <- decision_tree(
      tree_depth = current_params$tree_depth,
      min_n = current_params$min_n
      ) %>%
      set_engine('partykit',
      conditional_min_criterion = current_params$current_params
                 ) %>%
      set_mode('regression')
    
    # Fit model and collect metrics
    current_fit <- last_fit(current_spec, recipe, split = split)
    current_metrics <- current_fit %>% collect_metrics()
    
    # Extract RMSE value
    current_rmse <- current_metrics %>% 
      filter(.metric == "rmse") %>% 
      pull(.estimate)
    
    # Store results
    results <- results %>% add_row(
      tree_depth = current_params$tree_depth,
      min_n = current_params$min_n,
      conditional_min_criterion =current_params$conditional_min_criterion,
      rmse = current_rmse,
      source = current_params$source
    )
  }
  
  # Find the best combination
  best_params <- results %>% 
    arrange(rmse) %>% 
    slice(1)
  
  return(list(
    all_results = results,
    best_combination = best_params
  ))
}

# Example usage
results <- compare_hyperparameters(
  params_df = hyperparameters_df,
  recipe = weather_recipe,
  split = weather_split
)
print(results)
```




```{r final_spec}
final_spec <- decision_tree(tree_depth = best_rmse$tree_depth,
                            min_n = best_rmse$min_n
                            ) %>%
  # Specify the engine
  set_engine("partykit",
             conditional_min_criterion = best_rmse$conditional_min_criterion) %>%
  # Specifying mode  
  set_mode("regression")

final_spec
```

## 3. Validation  
```{r final_fit}
set.seed(10)
final_fit <- last_fit(final_spec,
                weather_recipe,
                split = weather_split)

final_fit %>%
  collect_predictions()
```
Metrics on the **test set**:
```{r}
final_fit %>%
  collect_metrics()
```
Metrics on **train set** (for curiosity and compare to test set):  
```{r}
# RMSE
final_spec %>%
  fit(grade_n ~ .,
      data = bake(weather_prep, 
                  weather_train)) %>%
  augment(new_data = bake(weather_prep, 
                          weather_train)) %>% 
  rmse(grade_n, .pred) %>%
  bind_rows(
    
    
    # R2
    final_spec %>%
      fit(grade_n ~ .,
          data = bake(weather_prep, 
                      weather_train)) %>%
      augment(new_data = bake(weather_prep, 
                              weather_train)) %>% 
      rsq(grade_n, .pred)
    
  )

```
```{r predicted vs. observed plot}
plot <- final_fit %>%
  collect_predictions() %>%
  metrica::scatter_plot(obs = grade_n,
                        pred = .pred,
                        print_eq = T,
                        print_metrics = T,
                        metrics_list = c("R2", "RMSE"),
                      # Customize metrics position
                        position_metrics = c(x=75,y=69),
                        position_eq = c(x=75, y =70.5))

plot
```


```{r export first plot}
ggsave(plot = plot,
       filename = "../output/ML_Grade/15_CART/S+O.png",
       width = 5,
       height = 5
       )
```

### Variable importance: 

```{r plotting VI}
set.seed(10)
plot_2 <- final_spec %>%
  fit(grade_n ~ .,
         data = bake(weather_prep, df)) %>%
    vip::vi() %>% 
  mutate(
    Importance = abs(Importance),
    Variable = fct_reorder(Variable, 
                           Importance)
  ) %>%
  slice_head(n = 10) %>%
  ggplot(aes(x = Importance, 
             y = Variable)) +
  geom_col() +
  scale_x_continuous(expand = c(0, 0)) +
  labs(y = NULL)

plot_2
```


```{r export second plot}
ggsave(plot = plot_2,
       filename = "../output/ML_Grade/15_CART/S+O_2.png",
       width = 5,
       height = 5,
       bg = "white"
       ) 
```

# 12) M
# Yield
```{r }
df <- read_sf("../data/data_complete.geojson") %>% 
  st_drop_geometry() %>% 
  as.data.frame() %>% 
  dplyr::select(planting_date, digger_date, growing_season, variety, river_basin, soil_series, row_pattern, 
                irrigation, saved_seed, 
                tillage_method=tillage_method_simplify,
                rotation = rotation_simplify_years_wo_peanut, seeding_rate, 
                grade_n, 
                alt, stage_duration_emergence:elevation,
                lat= lat.y, lon = long
                ) %>% 
  mutate_if(is.character, as.factor) %>% 
  mutate(rotation = as.factor(rotation)) %>% 
  dplyr::select(-(c(alt, rotation, seeding_rate, tillage_method))) %>% 
  dplyr::select(!(c(soil_alpha_0_5:soil_bd_30_60, 
                    soil_hb_0_5:soil_hb_30_60, 
                    soil_lambda_0_5:soil_n_30_60,
                    soil_theta_r_0_5:soil_theta_r_30_60
                    ))) %>%
  dplyr::select(-(c(SCI:NGRDI))) %>% 
  dplyr::select(grade_n, 
                planting_date:variety, 
                row_pattern:saved_seed, 
                lat, lon#, 
                #stage_duration_emergence:stage_vp_total,
                #river_basin, soil_series,
                #soil_clay_0_5:soil_theta_s_30_60,
                #NDVI:elevation
                ) %>% 
  #dplyr::select(-(c(NDVI:GNDVI, elevation))) %>% 
  drop_na() %>% dplyr::filter(grade_n >65) 

df
```
# ML workflow  
## 1. Pre-processing  
### a. Data split  
For data split, let's use **70% training / 30% testing**.
```{r weather_split}
# Setting seed to get reproducible results  
set.seed(27)
weather_split <- initial_split(df, prop = 0.7, 
                               strata = "grade_n")
weather_split
```
```{r weather_train}
# Setting train set 
weather_train <- training(weather_split)

weather_train
```


```{r weather_test}
# Setting test split
weather_test <- testing(weather_split)

weather_test
```
```{r data distribution}
# Combine the data frames
plot_train <- weather_train %>% mutate(Dataset = "Train")
plot_test <- weather_test %>% mutate(Dataset = "Test")
combined_data <- bind_rows(plot_train, plot_test)

# Create the density plot
ggplot(data = combined_data) +
  geom_density(aes(x = grade_n, fill = Dataset), alpha = 0.5) +
  scale_fill_manual(values = c("red", "blue")) +
  labs(x = "Yield (kg/ha)", y = "Density", fill = "Dataset")
```
  
### b. Data processing  
```{r weather_recipe}
weather_recipe <-
  # Defining predicted and predictor variables
  recipe(grade_n ~ .,
         data = weather_train) 
weather_recipe
```

```{r weather_prep}
weather_prep <- weather_recipe %>%
  prep()

weather_prep
```
## 2. Training  
### a. Model specification  
```{r hyperparameters}
cart_spec <- 
  decision_tree(tree_depth = tune(),
                min_n = tune()
                ) %>% 
  # Specify the engine
  set_engine("partykit", 
             conditional_min_criterion = tune()
             ) %>% 
  # Specifying mode  
  set_mode("regression")

cart_spec
```

### b. Hyper-parameter tuning  
  
```{r resampling_foldcv}
set.seed(15)
resampling_foldcv <- vfold_cv(weather_train, 
                              v = 5)

resampling_foldcv
```
```{r cit_param}
cit_param <- cart_spec %>% 
  extract_parameter_set_dials() %>% 
  update(conditional_min_criterion =conditional_min_criterion())

cit_param
cit_param$object
```

```{r rf_grid_result}
set.seed(45)
registerDoParallel(cores = parallel::detectCores() - 1)
rf_grid_result <- tune_race_anova(object = cart_spec,
                                   preprocessor = weather_recipe,
                                   resamples = resampling_foldcv,
                                   param_info = cit_param,
                                  grid = 50
                      )
stopImplicitCluster()
rf_grid_result$.metrics[[5]]
autoplot(rf_grid_result)
beepr::beep()
```

```{r hyperparameters selection by RMSE}
# Based on lowest RMSE
best_rmse <- rf_grid_result %>% 
  select_by_pct_loss("tree_depth",
                     metric = "rmse",
                     limit = 2
                     )%>% 
  mutate(source = "best_rmse")

best_rmse_2 <- rf_grid_result %>% 
  select_best(metric = "rmse")%>% 
  mutate(source = "best_rmse_2")

best_rmse_3 <- rf_grid_result %>% 
  select_by_one_std_err(metric = "rmse",
                        tree_depth
                        )%>% 
  mutate(source = "best_rmse_3")

```

```{r hyperparameters selection by R2}
# Based on greatest R2
best_r2 <- rf_grid_result %>% 
  select_by_pct_loss("tree_depth",
                     metric = "rsq",
                     limit = 2
                     ) %>% 
  mutate(source = "best_r2")

best_r2_2 <- rf_grid_result %>% 
  select_best(metric = "rsq")%>% 
  mutate(source = "best_r2_2")

best_r2_3 <- rf_grid_result %>% 
  select_by_one_std_err(metric = "rsq",
                        tree_depth
                        ) %>%
  mutate(source = "best_r2_3")
```
```{r comparing values}
hyperparameters_df <- best_rmse %>% 
  bind_rows(best_rmse_2, best_rmse_3, best_r2, best_r2_2, best_r2_3)
```

```{r}
# Function to find the best model
compare_hyperparameters <- function(params_df, recipe, split) {
  # Create empty tibble to store results
  results <- tibble(
    tree_depth = numeric(),
    min_n = numeric(),
    conditional_min_criterion = numeric(),
    rmse = numeric(),
    source = character()
  )
  
  # Loop through each row of parameters
  for(i in 1:nrow(params_df)) {
    # Extract current parameters
    current_params <- params_df[i, ]
    
    # Create model specification
    set.seed(10)
    current_spec <- decision_tree(
      tree_depth = current_params$tree_depth,
      min_n = current_params$min_n
      ) %>%
      set_engine('partykit',
      conditional_min_criterion = current_params$current_params
                 ) %>%
      set_mode('regression')
    
    # Fit model and collect metrics
    current_fit <- last_fit(current_spec, recipe, split = split)
    current_metrics <- current_fit %>% collect_metrics()
    
    # Extract RMSE value
    current_rmse <- current_metrics %>% 
      filter(.metric == "rmse") %>% 
      pull(.estimate)
    
    # Store results
    results <- results %>% add_row(
      tree_depth = current_params$tree_depth,
      min_n = current_params$min_n,
      conditional_min_criterion =current_params$conditional_min_criterion,
      rmse = current_rmse,
      source = current_params$source
    )
  }
  
  # Find the best combination
  best_params <- results %>% 
    arrange(rmse) %>% 
    slice(1)
  
  return(list(
    all_results = results,
    best_combination = best_params
  ))
}

# Example usage
results <- compare_hyperparameters(
  params_df = hyperparameters_df,
  recipe = weather_recipe,
  split = weather_split
)
print(results)
```




```{r final_spec}
final_spec <- decision_tree(tree_depth = best_rmse$tree_depth,
                            min_n = best_rmse$min_n
                            ) %>%
  # Specify the engine
  set_engine("partykit",
             conditional_min_criterion = best_rmse$conditional_min_criterion) %>%
  # Specifying mode  
  set_mode("regression")

final_spec
```

## 3. Validation  
```{r final_fit}
set.seed(10)
final_fit <- last_fit(final_spec,
                weather_recipe,
                split = weather_split)

final_fit %>%
  collect_predictions()
```
Metrics on the **test set**:
```{r}
final_fit %>%
  collect_metrics()
```
Metrics on **train set** (for curiosity and compare to test set):  
```{r}
# RMSE
final_spec %>%
  fit(grade_n ~ .,
      data = bake(weather_prep, 
                  weather_train)) %>%
  augment(new_data = bake(weather_prep, 
                          weather_train)) %>% 
  rmse(grade_n, .pred) %>%
  bind_rows(
    
    
    # R2
    final_spec %>%
      fit(grade_n ~ .,
          data = bake(weather_prep, 
                      weather_train)) %>%
      augment(new_data = bake(weather_prep, 
                              weather_train)) %>% 
      rsq(grade_n, .pred)
    
  )

```
```{r predicted vs. observed plot}
plot <- final_fit %>%
  collect_predictions() %>%
  metrica::scatter_plot(obs = grade_n,
                        pred = .pred,
                        print_eq = T,
                        print_metrics = T,
                        metrics_list = c("R2", "RMSE"),
                      # Customize metrics position
                        position_metrics = c(x=75,y=74),
                        position_eq = c(x=75, y =70.5))

plot
```


```{r export first plot}
ggsave(plot = plot,
       filename = "../output/ML_Grade/15_CART/M.png",
       width = 5,
       height = 5
       )
```

### Variable importance: 

```{r plotting VI}
set.seed(10)
plot_2 <- final_spec %>%
  fit(grade_n ~ .,
         data = bake(weather_prep, df)) %>%
    vip::vi() %>% 
  mutate(
    Importance = abs(Importance),
    Variable = fct_reorder(Variable, 
                           Importance)
  ) %>%
  slice_head(n = 10) %>%
  ggplot(aes(x = Importance, 
             y = Variable)) +
  geom_col() +
  scale_x_continuous(expand = c(0, 0)) +
  labs(y = NULL)

plot_2
```


```{r export second plot}
ggsave(plot = plot_2,
       filename = "../output/ML_Grade/15_CART/M_2.png",
       width = 5,
       height = 5,
       bg = "white"
       ) 
```

# 13) W
# Yield
```{r }
df <- read_sf("../data/data_complete.geojson") %>% 
  st_drop_geometry() %>% 
  as.data.frame() %>% 
  dplyr::select(planting_date, digger_date, growing_season, variety, river_basin, soil_series, row_pattern, 
                irrigation, saved_seed, 
                tillage_method=tillage_method_simplify,
                rotation = rotation_simplify_years_wo_peanut, seeding_rate, 
                grade_n, 
                alt, stage_duration_emergence:elevation,
                lat= lat.y, lon = long
                ) %>% 
  mutate_if(is.character, as.factor) %>% 
  mutate(rotation = as.factor(rotation)) %>% 
  dplyr::select(-(c(alt, rotation, seeding_rate, tillage_method))) %>% 
  dplyr::select(!(c(soil_alpha_0_5:soil_bd_30_60, 
                    soil_hb_0_5:soil_hb_30_60, 
                    soil_lambda_0_5:soil_n_30_60,
                    soil_theta_r_0_5:soil_theta_r_30_60
                    ))) %>%
  dplyr::select(-(c(SCI:NGRDI))) %>% 
  dplyr::select(grade_n, 
                #planting_date:variety, 
                #row_pattern:saved_seed, 
                #lat, lon, 
                stage_duration_emergence:stage_vp_total#,
                #river_basin, soil_series,
                #soil_clay_0_5:soil_theta_s_30_60,
                #NDVI:elevation
                ) %>% 
  #dplyr::select(-(c(NDVI:GNDVI, elevation))) %>% 
  drop_na() %>% dplyr::filter(grade_n >65) 

df
```
# ML workflow  
## 1. Pre-processing  
### a. Data split  
For data split, let's use **70% training / 30% testing**.
```{r weather_split}
# Setting seed to get reproducible results  
set.seed(27)
weather_split <- initial_split(df, prop = 0.7, 
                               strata = "grade_n")
weather_split
```
```{r weather_train}
# Setting train set 
weather_train <- training(weather_split)

weather_train
```


```{r weather_test}
# Setting test split
weather_test <- testing(weather_split)

weather_test
```
```{r data distribution}
# Combine the data frames
plot_train <- weather_train %>% mutate(Dataset = "Train")
plot_test <- weather_test %>% mutate(Dataset = "Test")
combined_data <- bind_rows(plot_train, plot_test)

# Create the density plot
ggplot(data = combined_data) +
  geom_density(aes(x = grade_n, fill = Dataset), alpha = 0.5) +
  scale_fill_manual(values = c("red", "blue")) +
  labs(x = "Yield (kg/ha)", y = "Density", fill = "Dataset")
```
  
### b. Data processing  
```{r weather_recipe}
weather_recipe <-
  # Defining predicted and predictor variables
  recipe(grade_n ~ .,
         data = weather_train) 
weather_recipe
```

```{r weather_prep}
weather_prep <- weather_recipe %>%
  prep()

weather_prep
```
## 2. Training  
### a. Model specification  
```{r hyperparameters}
cart_spec <- 
  decision_tree(tree_depth = tune(),
                min_n = tune()
                ) %>% 
  # Specify the engine
  set_engine("partykit", 
             conditional_min_criterion = tune()
             ) %>% 
  # Specifying mode  
  set_mode("regression")

cart_spec
```

### b. Hyper-parameter tuning  
  
```{r resampling_foldcv}
set.seed(15)
resampling_foldcv <- vfold_cv(weather_train, 
                              v = 5)

resampling_foldcv
```
```{r cit_param}
cit_param <- cart_spec %>% 
  extract_parameter_set_dials() %>% 
  update(conditional_min_criterion =conditional_min_criterion())

cit_param
cit_param$object
```

```{r rf_grid_result}
set.seed(45)
registerDoParallel(cores = parallel::detectCores() - 1)
rf_grid_result <- tune_race_anova(object = cart_spec,
                                   preprocessor = weather_recipe,
                                   resamples = resampling_foldcv,
                                   param_info = cit_param,
                                  grid = 50
                      )
stopImplicitCluster()
rf_grid_result$.metrics[[5]]
autoplot(rf_grid_result)
beepr::beep()
```

```{r hyperparameters selection by RMSE}
# Based on lowest RMSE
best_rmse <- rf_grid_result %>% 
  select_by_pct_loss("tree_depth",
                     metric = "rmse",
                     limit = 2
                     )%>% 
  mutate(source = "best_rmse")

best_rmse_2 <- rf_grid_result %>% 
  select_best(metric = "rmse")%>% 
  mutate(source = "best_rmse_2")

best_rmse_3 <- rf_grid_result %>% 
  select_by_one_std_err(metric = "rmse",
                        tree_depth
                        )%>% 
  mutate(source = "best_rmse_3")

```

```{r hyperparameters selection by R2}
# Based on greatest R2
best_r2 <- rf_grid_result %>% 
  select_by_pct_loss("tree_depth",
                     metric = "rsq",
                     limit = 2
                     ) %>% 
  mutate(source = "best_r2")

best_r2_2 <- rf_grid_result %>% 
  select_best(metric = "rsq")%>% 
  mutate(source = "best_r2_2")

best_r2_3 <- rf_grid_result %>% 
  select_by_one_std_err(metric = "rsq",
                        tree_depth
                        ) %>%
  mutate(source = "best_r2_3")
```
```{r comparing values}
hyperparameters_df <- best_rmse %>% 
  bind_rows(best_rmse_2, best_rmse_3)
```

```{r}
# Function to find the best model
compare_hyperparameters <- function(params_df, recipe, split) {
  # Create empty tibble to store results
  results <- tibble(
    tree_depth = numeric(),
    min_n = numeric(),
    conditional_min_criterion = numeric(),
    rmse = numeric(),
    source = character()
  )
  
  # Loop through each row of parameters
  for(i in 1:nrow(params_df)) {
    # Extract current parameters
    current_params <- params_df[i, ]
    
    # Create model specification
    set.seed(10)
    current_spec <- decision_tree(
      tree_depth = current_params$tree_depth,
      min_n = current_params$min_n
      ) %>%
      set_engine('partykit',
      conditional_min_criterion = current_params$current_params
                 ) %>%
      set_mode('regression')
    
    # Fit model and collect metrics
    current_fit <- last_fit(current_spec, recipe, split = split)
    current_metrics <- current_fit %>% collect_metrics()
    
    # Extract RMSE value
    current_rmse <- current_metrics %>% 
      filter(.metric == "rmse") %>% 
      pull(.estimate)
    
    # Store results
    results <- results %>% add_row(
      tree_depth = current_params$tree_depth,
      min_n = current_params$min_n,
      conditional_min_criterion =current_params$conditional_min_criterion,
      rmse = current_rmse,
      source = current_params$source
    )
  }
  
  # Find the best combination
  best_params <- results %>% 
    arrange(rmse) %>% 
    slice(1)
  
  return(list(
    all_results = results,
    best_combination = best_params
  ))
}

# Example usage
results <- compare_hyperparameters(
  params_df = hyperparameters_df,
  recipe = weather_recipe,
  split = weather_split
)
print(results)
```




```{r final_spec}
final_spec <- decision_tree(tree_depth = best_rmse$tree_depth,
                            min_n = best_rmse$min_n
                            ) %>%
  # Specify the engine
  set_engine("partykit",
             conditional_min_criterion = best_rmse$conditional_min_criterion) %>%
  # Specifying mode  
  set_mode("regression")

final_spec
```

## 3. Validation  
```{r final_fit}
set.seed(10)
final_fit <- last_fit(final_spec,
                weather_recipe,
                split = weather_split)

final_fit %>%
  collect_predictions()
```
Metrics on the **test set**:
```{r}
final_fit %>%
  collect_metrics()
```
Metrics on **train set** (for curiosity and compare to test set):  
```{r}
# RMSE
final_spec %>%
  fit(grade_n ~ .,
      data = bake(weather_prep, 
                  weather_train)) %>%
  augment(new_data = bake(weather_prep, 
                          weather_train)) %>% 
  rmse(grade_n, .pred) %>%
  bind_rows(
    
    
    # R2
    final_spec %>%
      fit(grade_n ~ .,
          data = bake(weather_prep, 
                      weather_train)) %>%
      augment(new_data = bake(weather_prep, 
                              weather_train)) %>% 
      rsq(grade_n, .pred)
    
  )

```
```{r predicted vs. observed plot}
plot <- final_fit %>%
  collect_predictions() %>%
  metrica::scatter_plot(obs = grade_n,
                        pred = .pred,
                        print_eq = T,
                        print_metrics = T,
                        metrics_list = c("R2", "RMSE"),
                      # Customize metrics position
                        position_metrics = c(x=75,y=73),
                        position_eq = c(x=75, y =69))

plot
```


```{r export first plot}
ggsave(plot = plot,
       filename = "../output/ML_Grade/15_CART/W.png",
       width = 5,
       height = 5
       )
```

### Variable importance: 

```{r plotting VI}
set.seed(10)
plot_2 <- final_spec %>%
  fit(grade_n ~ .,
         data = bake(weather_prep, df)) %>%
    vip::vi() %>% 
  mutate(
    Importance = abs(Importance),
    Variable = fct_reorder(Variable, 
                           Importance)
  ) %>%
  slice_head(n = 10) %>%
  ggplot(aes(x = Importance, 
             y = Variable)) +
  geom_col() +
  scale_x_continuous(expand = c(0, 0)) +
  labs(y = NULL)

plot_2
```


```{r export second plot}
ggsave(plot = plot_2,
       filename = "../output/ML_Grade/15_CART/W_2.png",
       width = 5,
       height = 5,
       bg = "white"
       ) 
```

# 13) S
# Yield
```{r }
df <- read_sf("../data/data_complete.geojson") %>% 
  st_drop_geometry() %>% 
  as.data.frame() %>% 
  dplyr::select(planting_date, digger_date, growing_season, variety, river_basin, soil_series, row_pattern, 
                irrigation, saved_seed, 
                tillage_method=tillage_method_simplify,
                rotation = rotation_simplify_years_wo_peanut, seeding_rate, 
                grade_n, 
                alt, stage_duration_emergence:elevation,
                lat= lat.y, lon = long
                ) %>% 
  mutate_if(is.character, as.factor) %>% 
  mutate(rotation = as.factor(rotation)) %>% 
  dplyr::select(-(c(alt, rotation, seeding_rate, tillage_method))) %>% 
  dplyr::select(!(c(soil_alpha_0_5:soil_bd_30_60, 
                    soil_hb_0_5:soil_hb_30_60, 
                    soil_lambda_0_5:soil_n_30_60,
                    soil_theta_r_0_5:soil_theta_r_30_60
                    ))) %>%
  dplyr::select(-(c(SCI:NGRDI))) %>% 
  dplyr::select(grade_n, 
                #planting_date:variety, 
                #row_pattern:saved_seed, 
                #lat, lon, 
                #stage_duration_emergence:stage_vp_total,
                river_basin, soil_series,
                soil_clay_0_5:soil_theta_s_30_60#,
                #NDVI:elevation
                ) %>% 
  #dplyr::select(-(c(NDVI:GNDVI, elevation))) %>% 
  drop_na() %>% dplyr::filter(grade_n >65) 

df
```
# ML workflow  
## 1. Pre-processing  
### a. Data split  
For data split, let's use **70% training / 30% testing**.
```{r weather_split}
# Setting seed to get reproducible results  
set.seed(27)
weather_split <- initial_split(df, prop = 0.7, 
                               strata = "grade_n")
weather_split
```
```{r weather_train}
# Setting train set 
weather_train <- training(weather_split)

weather_train
```


```{r weather_test}
# Setting test split
weather_test <- testing(weather_split)

weather_test
```
```{r data distribution}
# Combine the data frames
plot_train <- weather_train %>% mutate(Dataset = "Train")
plot_test <- weather_test %>% mutate(Dataset = "Test")
combined_data <- bind_rows(plot_train, plot_test)

# Create the density plot
ggplot(data = combined_data) +
  geom_density(aes(x = grade_n, fill = Dataset), alpha = 0.5) +
  scale_fill_manual(values = c("red", "blue")) +
  labs(x = "Yield (kg/ha)", y = "Density", fill = "Dataset")
```
  
### b. Data processing  
```{r weather_recipe}
weather_recipe <-
  # Defining predicted and predictor variables
  recipe(grade_n ~ .,
         data = weather_train) 
weather_recipe
```

```{r weather_prep}
weather_prep <- weather_recipe %>%
  prep()

weather_prep
```
## 2. Training  
### a. Model specification  
```{r hyperparameters}
cart_spec <- 
  decision_tree(tree_depth = tune(),
                min_n = tune()
                ) %>% 
  # Specify the engine
  set_engine("partykit", 
             conditional_min_criterion = tune()
             ) %>% 
  # Specifying mode  
  set_mode("regression")

cart_spec
```

### b. Hyper-parameter tuning  
  
```{r resampling_foldcv}
set.seed(15)
resampling_foldcv <- vfold_cv(weather_train, 
                              v = 5)

resampling_foldcv
```
```{r cit_param}
cit_param <- cart_spec %>% 
  extract_parameter_set_dials() %>% 
  update(conditional_min_criterion =conditional_min_criterion())

cit_param
cit_param$object
```

```{r rf_grid_result}
set.seed(45)
registerDoParallel(cores = parallel::detectCores() - 1)
rf_grid_result <- tune_race_anova(object = cart_spec,
                                   preprocessor = weather_recipe,
                                   resamples = resampling_foldcv,
                                   param_info = cit_param,
                                  grid = 50
                      )
stopImplicitCluster()
rf_grid_result$.metrics[[5]]
autoplot(rf_grid_result)
beepr::beep()
```

```{r hyperparameters selection by RMSE}
# Based on lowest RMSE
best_rmse <- rf_grid_result %>% 
  select_by_pct_loss("tree_depth",
                     metric = "rmse",
                     limit = 2
                     )%>% 
  mutate(source = "best_rmse")

best_rmse_2 <- rf_grid_result %>% 
  select_best(metric = "rmse")%>% 
  mutate(source = "best_rmse_2")

best_rmse_3 <- rf_grid_result %>% 
  select_by_one_std_err(metric = "rmse",
                        tree_depth
                        )%>% 
  mutate(source = "best_rmse_3")

```

```{r hyperparameters selection by R2}
# Based on greatest R2
best_r2 <- rf_grid_result %>% 
  select_by_pct_loss("tree_depth",
                     metric = "rsq",
                     limit = 2
                     ) %>% 
  mutate(source = "best_r2")

best_r2_2 <- rf_grid_result %>% 
  select_best(metric = "rsq")%>% 
  mutate(source = "best_r2_2")

best_r2_3 <- rf_grid_result %>% 
  select_by_one_std_err(metric = "rsq",
                        tree_depth
                        ) %>%
  mutate(source = "best_r2_3")
```
```{r comparing values}
hyperparameters_df <- best_rmse %>% 
  bind_rows(best_rmse_2, best_rmse_3, best_r2, best_r2_2, best_r2_3)
```

```{r}
# Function to find the best model
compare_hyperparameters <- function(params_df, recipe, split) {
  # Create empty tibble to store results
  results <- tibble(
    tree_depth = numeric(),
    min_n = numeric(),
    conditional_min_criterion = numeric(),
    rmse = numeric(),
    source = character()
  )
  
  # Loop through each row of parameters
  for(i in 1:nrow(params_df)) {
    # Extract current parameters
    current_params <- params_df[i, ]
    
    # Create model specification
    set.seed(10)
    current_spec <- decision_tree(
      tree_depth = current_params$tree_depth,
      min_n = current_params$min_n
      ) %>%
      set_engine('partykit',
      conditional_min_criterion = current_params$current_params
                 ) %>%
      set_mode('regression')
    
    # Fit model and collect metrics
    current_fit <- last_fit(current_spec, recipe, split = split)
    current_metrics <- current_fit %>% collect_metrics()
    
    # Extract RMSE value
    current_rmse <- current_metrics %>% 
      filter(.metric == "rmse") %>% 
      pull(.estimate)
    
    # Store results
    results <- results %>% add_row(
      tree_depth = current_params$tree_depth,
      min_n = current_params$min_n,
      conditional_min_criterion =current_params$conditional_min_criterion,
      rmse = current_rmse,
      source = current_params$source
    )
  }
  
  # Find the best combination
  best_params <- results %>% 
    arrange(rmse) %>% 
    slice(1)
  
  return(list(
    all_results = results,
    best_combination = best_params
  ))
}

# Example usage
results <- compare_hyperparameters(
  params_df = hyperparameters_df,
  recipe = weather_recipe,
  split = weather_split
)
print(results)
```




```{r final_spec}
final_spec <- decision_tree(tree_depth = best_rmse$tree_depth,
                            min_n = best_rmse$min_n
                            ) %>%
  # Specify the engine
  set_engine("partykit",
             conditional_min_criterion = best_rmse$conditional_min_criterion) %>%
  # Specifying mode  
  set_mode("regression")

final_spec
```

## 3. Validation  
```{r final_fit}
set.seed(10)
final_fit <- last_fit(final_spec,
                weather_recipe,
                split = weather_split)

final_fit %>%
  collect_predictions()
```
Metrics on the **test set**:
```{r}
final_fit %>%
  collect_metrics()
```
Metrics on **train set** (for curiosity and compare to test set):  
```{r}
# RMSE
final_spec %>%
  fit(grade_n ~ .,
      data = bake(weather_prep, 
                  weather_train)) %>%
  augment(new_data = bake(weather_prep, 
                          weather_train)) %>% 
  rmse(grade_n, .pred) %>%
  bind_rows(
    
    
    # R2
    final_spec %>%
      fit(grade_n ~ .,
          data = bake(weather_prep, 
                      weather_train)) %>%
      augment(new_data = bake(weather_prep, 
                              weather_train)) %>% 
      rsq(grade_n, .pred)
    
  )

```
```{r predicted vs. observed plot}
plot <- final_fit %>%
  collect_predictions() %>%
  metrica::scatter_plot(obs = grade_n,
                        pred = .pred,
                        print_eq = T,
                        print_metrics = T,
                        metrics_list = c("R2", "RMSE"),
                      # Customize metrics position
                        position_metrics = c(x=75,y=73),
                        position_eq = c(x=75, y =70.5))

plot
```


```{r export first plot}
ggsave(plot = plot,
       filename = "../output/ML_Grade/15_CART/S.png",
       width = 5,
       height = 5
       )
```

### Variable importance: 

```{r plotting VI}
set.seed(10)
plot_2 <- final_spec %>%
  fit(grade_n ~ .,
         data = bake(weather_prep, df)) %>%
    vip::vi() %>% 
  mutate(
    Importance = abs(Importance),
    Variable = fct_reorder(Variable, 
                           Importance)
  ) %>%
  slice_head(n = 10) %>%
  ggplot(aes(x = Importance, 
             y = Variable)) +
  geom_col() +
  scale_x_continuous(expand = c(0, 0)) +
  labs(y = NULL)

plot_2
```


```{r export second plot}
ggsave(plot = plot_2,
       filename = "../output/ML_Grade/15_CART/S_2.png",
       width = 5,
       height = 5,
       bg = "white"
       ) 
```

# 15) O
# Yield
```{r }
df <- read_sf("../data/data_complete.geojson") %>% 
  st_drop_geometry() %>% 
  as.data.frame() %>% 
  dplyr::select(planting_date, digger_date, growing_season, variety, river_basin, soil_series, row_pattern, 
                irrigation, saved_seed, 
                tillage_method=tillage_method_simplify,
                rotation = rotation_simplify_years_wo_peanut, seeding_rate, 
                grade_n, 
                alt, stage_duration_emergence:elevation,
                lat= lat.y, lon = long
                ) %>% 
  mutate_if(is.character, as.factor) %>% 
  mutate(rotation = as.factor(rotation)) %>% 
  dplyr::select(-(c(alt, rotation, seeding_rate, tillage_method))) %>% 
  dplyr::select(!(c(soil_alpha_0_5:soil_bd_30_60, 
                    soil_hb_0_5:soil_hb_30_60, 
                    soil_lambda_0_5:soil_n_30_60,
                    soil_theta_r_0_5:soil_theta_r_30_60
                    ))) %>%
  dplyr::select(-(c(SCI:NGRDI))) %>% 
  dplyr::select(grade_n, 
                #planting_date:variety, 
                #row_pattern:saved_seed, 
                #lat, lon, 
                #stage_duration_emergence:stage_vp_total,
                #river_basin, soil_series,
                #soil_clay_0_5:soil_theta_s_30_60,
                NDVI:elevation
                ) %>% 
  #dplyr::select(-(c(NDVI:GNDVI, elevation))) %>% 
  drop_na() %>% dplyr::filter(grade_n >65) 

df
```
# ML workflow  
## 1. Pre-processing  
### a. Data split  
For data split, let's use **70% training / 30% testing**.
```{r weather_split}
# Setting seed to get reproducible results  
set.seed(27)
weather_split <- initial_split(df, prop = 0.7, 
                               strata = "grade_n")
weather_split
```
```{r weather_train}
# Setting train set 
weather_train <- training(weather_split)

weather_train
```


```{r weather_test}
# Setting test split
weather_test <- testing(weather_split)

weather_test
```
```{r data distribution}
# Combine the data frames
plot_train <- weather_train %>% mutate(Dataset = "Train")
plot_test <- weather_test %>% mutate(Dataset = "Test")
combined_data <- bind_rows(plot_train, plot_test)

# Create the density plot
ggplot(data = combined_data) +
  geom_density(aes(x = grade_n, fill = Dataset), alpha = 0.5) +
  scale_fill_manual(values = c("red", "blue")) +
  labs(x = "Yield (kg/ha)", y = "Density", fill = "Dataset")
```
  
### b. Data processing  
```{r weather_recipe}
weather_recipe <-
  # Defining predicted and predictor variables
  recipe(grade_n ~ .,
         data = weather_train) 
weather_recipe
```

```{r weather_prep}
weather_prep <- weather_recipe %>%
  prep()

weather_prep
```
## 2. Training  
### a. Model specification  
```{r hyperparameters}
cart_spec <- 
  decision_tree(tree_depth = tune(),
                min_n = tune()
                ) %>% 
  # Specify the engine
  set_engine("partykit", 
             conditional_min_criterion = tune()
             ) %>% 
  # Specifying mode  
  set_mode("regression")

cart_spec
```

### b. Hyper-parameter tuning  
  
```{r resampling_foldcv}
set.seed(15)
resampling_foldcv <- vfold_cv(weather_train, 
                              v = 5)

resampling_foldcv
```
```{r cit_param}
cit_param <- cart_spec %>% 
  extract_parameter_set_dials() %>% 
  update(conditional_min_criterion =conditional_min_criterion())

cit_param
cit_param$object
```

```{r rf_grid_result}
set.seed(45)
registerDoParallel(cores = parallel::detectCores() - 1)
rf_grid_result <- tune_race_anova(object = cart_spec,
                                   preprocessor = weather_recipe,
                                   resamples = resampling_foldcv,
                                   param_info = cit_param,
                                  grid = 50
                      )
stopImplicitCluster()
rf_grid_result$.metrics[[5]]
autoplot(rf_grid_result)
beepr::beep()
```

```{r hyperparameters selection by RMSE}
# Based on lowest RMSE
best_rmse <- rf_grid_result %>% 
  select_by_pct_loss("tree_depth",
                     metric = "rmse",
                     limit = 2
                     )%>% 
  mutate(source = "best_rmse")

best_rmse_2 <- rf_grid_result %>% 
  select_best(metric = "rmse")%>% 
  mutate(source = "best_rmse_2")

best_rmse_3 <- rf_grid_result %>% 
  select_by_one_std_err(metric = "rmse",
                        tree_depth
                        )%>% 
  mutate(source = "best_rmse_3")

```

```{r hyperparameters selection by R2}
# Based on greatest R2
best_r2 <- rf_grid_result %>% 
  select_by_pct_loss("tree_depth",
                     metric = "rsq",
                     limit = 2
                     ) %>% 
  mutate(source = "best_r2")

best_r2_2 <- rf_grid_result %>% 
  select_best(metric = "rsq")%>% 
  mutate(source = "best_r2_2")

best_r2_3 <- rf_grid_result %>% 
  select_by_one_std_err(metric = "rsq",
                        tree_depth
                        ) %>%
  mutate(source = "best_r2_3")
```
```{r comparing values}
hyperparameters_df <- best_rmse %>% 
  bind_rows(best_rmse_2, best_rmse_3)
```

```{r}
# Function to find the best model
compare_hyperparameters <- function(params_df, recipe, split) {
  # Create empty tibble to store results
  results <- tibble(
    tree_depth = numeric(),
    min_n = numeric(),
    conditional_min_criterion = numeric(),
    rmse = numeric(),
    source = character()
  )
  
  # Loop through each row of parameters
  for(i in 1:nrow(params_df)) {
    # Extract current parameters
    current_params <- params_df[i, ]
    
    # Create model specification
    set.seed(10)
    current_spec <- decision_tree(
      tree_depth = current_params$tree_depth,
      min_n = current_params$min_n
      ) %>%
      set_engine('partykit',
      conditional_min_criterion = current_params$current_params
                 ) %>%
      set_mode('regression')
    
    # Fit model and collect metrics
    current_fit <- last_fit(current_spec, recipe, split = split)
    current_metrics <- current_fit %>% collect_metrics()
    
    # Extract RMSE value
    current_rmse <- current_metrics %>% 
      filter(.metric == "rmse") %>% 
      pull(.estimate)
    
    # Store results
    results <- results %>% add_row(
      tree_depth = current_params$tree_depth,
      min_n = current_params$min_n,
      conditional_min_criterion =current_params$conditional_min_criterion,
      rmse = current_rmse,
      source = current_params$source
    )
  }
  
  # Find the best combination
  best_params <- results %>% 
    arrange(rmse) %>% 
    slice(1)
  
  return(list(
    all_results = results,
    best_combination = best_params
  ))
}

# Example usage
results <- compare_hyperparameters(
  params_df = hyperparameters_df,
  recipe = weather_recipe,
  split = weather_split
)
print(results)
```




```{r final_spec}
final_spec <- decision_tree(tree_depth = best_rmse$tree_depth,
                            min_n = best_rmse$min_n
                            ) %>%
  # Specify the engine
  set_engine("partykit",
             conditional_min_criterion = best_rmse$conditional_min_criterion) %>%
  # Specifying mode  
  set_mode("regression")

final_spec
```

## 3. Validation  
```{r final_fit}
set.seed(10)
final_fit <- last_fit(final_spec,
                weather_recipe,
                split = weather_split)

final_fit %>%
  collect_predictions()
```
Metrics on the **test set**:
```{r}
final_fit %>%
  collect_metrics()
```
Metrics on **train set** (for curiosity and compare to test set):  
```{r}
# RMSE
final_spec %>%
  fit(grade_n ~ .,
      data = bake(weather_prep, 
                  weather_train)) %>%
  augment(new_data = bake(weather_prep, 
                          weather_train)) %>% 
  rmse(grade_n, .pred) %>%
  bind_rows(
    
    
    # R2
    final_spec %>%
      fit(grade_n ~ .,
          data = bake(weather_prep, 
                      weather_train)) %>%
      augment(new_data = bake(weather_prep, 
                              weather_train)) %>% 
      rsq(grade_n, .pred)
    
  )

```
```{r predicted vs. observed plot}
plot <- final_fit %>%
  collect_predictions() %>%
  metrica::scatter_plot(obs = grade_n,
                        pred = .pred,
                        print_eq = T,
                        print_metrics = T,
                        metrics_list = c("R2", "RMSE"),
                      # Customize metrics position
                        position_metrics = c(x=75,y=73),
                        position_eq = c(x=75, y =71))

plot
```


```{r export first plot}
ggsave(plot = plot,
       filename = "../output/ML_Grade/15_CART/O.png",
       width = 5,
       height = 5
       )
```

### Variable importance: 

```{r plotting VI}
set.seed(10)
plot_2 <- final_spec %>%
  fit(grade_n ~ .,
         data = bake(weather_prep, df)) %>%
    vip::vi() %>% 
  mutate(
    Importance = abs(Importance),
    Variable = fct_reorder(Variable, 
                           Importance)
  ) %>%
  slice_head(n = 10) %>%
  ggplot(aes(x = Importance, 
             y = Variable)) +
  geom_col() +
  scale_x_continuous(expand = c(0, 0)) +
  labs(y = NULL)

plot_2
```


```{r export second plot}
ggsave(plot = plot_2,
       filename = "../output/ML_Grade/15_CART/O_2.png",
       width = 5,
       height = 5,
       bg = "white"
       ) 
```

